---
title: "Statistical Analysis: Win Rates"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom.mixed)

source("favorability_test.R")
source("data_helpers.R")
```

# Load data

Vendor identities are blinded by `blind_vendors.R`, which writes the
`blinded_*.csv` files read below and a `secret_mapping.csv` file (kept
outside the analysis pipeline). Run
`Rscript blind_vendors.R pilot_data secret/secret_mapping.csv`
before knitting this document.

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
users <- get_eligible_users(users)
prior_users_path <- file.path(data_dir, "blinded_prior_users.csv")
prior_users <- if (file.exists(prior_users_path)) {
  read_csv(prior_users_path)
} else {
  NULL
}
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users, prior_users = prior_users)
qa_ratings <- apply_inclusion_criteria(qa_ratings)
```

# Primary Analysis: Favorability Score Tests

One-vs-rest favorability score test for each vendor on the two primary
outcome axes (accuracy and clinical_utility), restricted to the text-only
rendering condition. Bootstrap resamples questions as clusters.

```{r favorability_test_helpers}
prepare_one_vs_rest_df <- function(df, vendor) {
  df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      outcome = case_when(
        winner == vendor ~ 1L,
        loser == vendor ~ -1L,
        is_tie ~ 0L
      )
    ) %>%
    select(question_id, outcome)
}

run_favorability_analysis <- function(df, axis_name,
                                      subversion_filter = "qa_text_only",
                                      n_boot = 2000, seed = 42) {
  set.seed(seed)

  axis_df <- df %>%
    filter(subversion == subversion_filter, axis == axis_name)

  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    test_df <- prepare_one_vs_rest_df(axis_df, v)

    fav_score <- get_favorability_score(test_df$outcome)
    test_result <- run_favorability_test(test_df, n_boot = n_boot, ci_level=0.95)

    tibble(
      vendor = v,
      wins = sum(test_df$outcome == 1),
      ties = sum(test_df$outcome == 0),
      losses = sum(test_df$outcome == -1),
      n = nrow(test_df),
      favorability = fav_score,
      ci_lower = test_result$ci_lower,
      ci_upper = test_result$ci_upper,
      p_value = test_result$p_value
    )
  })
}

format_favorability_table <- function(results, axis_name) {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
      p_value = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
    kable(caption = sprintf(
      "One-vs-rest favorability score test (text-only, %s axis)",
      axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility

```{r favorability_test_clinical}
run_favorability_analysis(qa_ratings, "clinical_utility") %>%
  format_favorability_table("clinical utility")
```

## Accuracy

```{r favorability_test_accuracy}
run_favorability_analysis(qa_ratings, "accuracy") %>%
  format_favorability_table("accuracy")
```

## Authoritativeness

```{r favorability_test_authoritativeness}
run_favorability_analysis(qa_ratings, "authoritativeness") %>%
  format_favorability_table("authoritativeness")
```

## Completeness

```{r favorability_test_completeness}
run_favorability_analysis(qa_ratings, "completeness") %>%
  format_favorability_table("completeness")
```

## Verifiability

```{r favorability_test_verifiability}
run_favorability_analysis(qa_ratings, "verifiability") %>%
  format_favorability_table("verifiability")
```

# Secondary analysis: Win-rate tables

For every (assignment, axis) row we know which two vendors were compared
and the chosen winner. The matrix below reports, for each ordered pair
(row vendor vs. column vendor), the row vendor's win rate across all
axes. Ties are *included* in the denominator (and count as 0 wins for
both vendors), so cell[row, col] + cell[col, row] + tie-rate = 100%.

These matrices, the win-margin tables below, and the Plackett-Luce
rankings are all restricted to the `qa_text_only` and
`qa_text_citations` rendering modes (the `qa_full_render` condition is
excluded).

```{r included_subversions}
included_subversions <- c("qa_text_only", "qa_text_citations")
qa_ratings_included <- qa_ratings %>%
  filter(subversion %in% included_subversions)
```

## Win-rate matrix (aggregate)

```{r winrate_matrix}
build_win_matrix <- function(df, n_boot = 1000) {
  df %>%
    mutate(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_win = !is_tie & winner == slot_a_vendor
    ) %>%
    bind_rows(
      df %>%
        mutate(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_win = !is_tie & winner == slot_b_vendor
        )
    ) %>%
    select(question_id, row_vendor, col_vendor, row_win) %>%
    group_by(row_vendor, col_vendor) %>%
    group_modify(~ {
      res <- run_win_rate_ci(
        data.frame(question_id = .x$question_id, row_win = .x$row_win),
        n_boot = n_boot
      )
      tibble(
        n        = res$n,
        wins     = res$wins,
        win_rate = res$win_rate,
        ci_lower = res$ci_lower,
        ci_upper = res$ci_upper
      )
    }) %>%
    ungroup() %>%
    mutate(cell = sprintf("%.1f%% [%.1f, %.1f] (%d/%d)",
                          100 * win_rate,
                          100 * ci_lower,
                          100 * ci_upper,
                          wins, n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
                values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}

set.seed(42)

kable(
  build_win_matrix(qa_ratings_included),
  caption = paste0(
    "Row vendor's win rate vs. column vendor (ties included in ",
    "denominator). Cell = win% [95% CI] (wins / all comparisons). ",
    "CI from question-clustered bootstrap (1000 resamples)."
  )
) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Win-rate matrices per axis

Same construction as the pooled matrix, repeated within each rating axis.

```{r winrate_matrix_per_axis, results='asis'}
set.seed(42)

axes <- sort(unique(qa_ratings$axis))

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings_included %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor on axis = %s ",
               "(ties included). Cell = win%% [95%% CI] (wins/N)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

## Win-rate matrices per rendering mode

Same construction as the pooled matrix, repeated within each of the
two included rendering modes (`subversion` in `users.csv`).

```{r winrate_matrix_per_subversion, results='asis'}
for (sv in included_subversions) {
  cat(sprintf("\n\n### Rendering mode: %s\n\n", sv))
  mat <- build_win_matrix(qa_ratings_included %>% filter(subversion == sv))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor for ",
               "rendering mode = %s (ties included in denominator)."),
        sv
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

# Sensitivity analysis: Win-margin tables

Same row-vs-column structure as the win-rate matrices, but each cell is
the mean number of Likert units the row vendor beat the column vendor
by. The Likert scale is `strongly_*` = 2 units, `slightly_*` = 1 unit,
`tie` = 0 units. Ties are *included* (they contribute 0 units). Cells
are mirrored: `cell[row, col] = -cell[col, row]`. Cell format:
`mean (net/N)` where `net` is the signed sum of units across all
comparisons in that cell.

```{r margin_helpers}
likert_units_a <- function(choice) {
  case_when(
    choice == "strongly_a" ~  2,
    choice == "slightly_a" ~  1,
    choice == "tie"        ~  0,
    choice == "slightly_b" ~ -1,
    choice == "strongly_b" ~ -2
  )
}

build_margin_matrix <- function(df) {
  df %>%
    mutate(units_a = likert_units_a(choice)) %>%
    transmute(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_units  = units_a
    ) %>%
    bind_rows(
      df %>%
        mutate(units_a = likert_units_a(choice)) %>%
        transmute(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_units  = -units_a
        )
    ) %>%
    group_by(row_vendor, col_vendor) %>%
    summarise(
      n = n(),
      net_units  = sum(row_units),
      mean_units = net_units / n,
      .groups = "drop"
    ) %>%
    mutate(cell = sprintf("%+.2f (%+d/%d)",
                          mean_units,
                          as.integer(net_units),
                          n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
                values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}
```

## Win-margin matrix

```{r margin_matrix}
kable(
  build_margin_matrix(qa_ratings_included),
  caption = paste0(
    "Mean Likert-unit margin the row vendor won by vs. the column ",
    "vendor (ties included as 0). Cell = mean (net/N)."
  )
) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "hover"))
```

## Win-margin matrices per axis

```{r margin_matrix_per_axis, results='asis'}
for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_margin_matrix(qa_ratings_included %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Mean Likert-unit margin row vendor won by vs. column ",
               "vendor on axis = %s (ties included)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

## Win-margin matrices per rendering mode

```{r margin_matrix_per_subversion, results='asis'}
for (sv in included_subversions) {
  cat(sprintf("\n\n### Rendering mode: %s\n\n", sv))
  mat <- build_margin_matrix(qa_ratings_included %>% filter(subversion == sv))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Mean Likert-unit margin row vendor won by vs. column ",
               "vendor for rendering mode = %s (ties included)."),
        sv
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

# Sensitivity analysis: Length-stratified favorability tests

Re-runs the one-vs-rest favorability score test, but stratified by
whether the vendor under test produced a *longer* or *shorter* answer
than its comparator (character length from `blinded_qa_answers.csv`).
Unlike the primary analysis, rendering conditions are **pooled** here
(all `subversion` values included). The comparison count drops because
each comparison goes into exactly one stratum; equal-length answers are
not expected and raise an error if encountered.

```{r length_stratified_helpers}
answer_lengths <- read_csv(file.path(data_dir, "blinded_qa_answers.csv")) %>%
  select(question_id, vendor, answer_len)

qa_ratings_with_lengths <- qa_ratings %>%
  left_join(
    answer_lengths %>%
      rename(slot_a_vendor = vendor, slot_a_answer_len = answer_len),
    by = c("question_id", "slot_a_vendor")
  ) %>%
  left_join(
    answer_lengths %>%
      rename(slot_b_vendor = vendor, slot_b_answer_len = answer_len),
    by = c("question_id", "slot_b_vendor")
  )

prepare_one_vs_rest_df_with_lengths <- function(df, vendor) {
  df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      vendor_len     = if_else(slot_a_vendor == vendor,
                               slot_a_answer_len, slot_b_answer_len),
      comparator_len = if_else(slot_a_vendor == vendor,
                               slot_b_answer_len, slot_a_answer_len),
      outcome = case_when(
        winner == vendor ~ 1L,
        loser  == vendor ~ -1L,
        is_tie ~ 0L
      )
    ) %>%
    select(question_id, outcome, vendor_len, comparator_len)
}

run_length_stratified_analysis <- function(df, axis_name,
                                           n_boot = 2000, seed = 42) {
  set.seed(seed)

  axis_df <- df %>% filter(axis == axis_name)
  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    base_df <- prepare_one_vs_rest_df_with_lengths(axis_df, v)

    if (any(base_df$vendor_len == base_df$comparator_len)) {
      stop(sprintf(
        "Equal answer lengths for vendor %s on axis %s — unexpected.",
        v, axis_name
      ))
    }

    map_dfr(c("longer", "shorter"), function(stratum) {
      stratum_df <- base_df %>%
        filter(if (stratum == "longer") vendor_len > comparator_len
                                  else vendor_len < comparator_len) %>%
        select(question_id, outcome)

      fav_score   <- get_favorability_score(stratum_df$outcome)
      test_result <- run_favorability_test(stratum_df,
                                           n_boot = n_boot, ci_level = 0.95)

      tibble(
        vendor       = v,
        stratum      = stratum,
        wins         = sum(stratum_df$outcome == 1),
        ties         = sum(stratum_df$outcome == 0),
        losses       = sum(stratum_df$outcome == -1),
        n            = nrow(stratum_df),
        favorability = fav_score,
        ci_lower     = test_result$ci_lower,
        ci_upper     = test_result$ci_upper,
        p_value      = test_result$p_value
      )
    })
  })
}

format_length_stratified_table <- function(results, axis_name) {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95        = sprintf("[%.1f%%, %.1f%%]",
                             100 * ci_lower, 100 * ci_upper),
      p_value      = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, stratum, wins, ties, losses, n,
           favorability, ci_95, p_value) %>%
    arrange(vendor, stratum) %>%
    kable(caption = sprintf(
      paste0("Length-stratified one-vs-rest favorability test ",
             "(pooled rendering, %s axis). Stratum = vendor's answer ",
             "longer/shorter than comparator."),
      axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility

```{r length_stratified_clinical}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "clinical_utility") %>%
  format_length_stratified_table("clinical utility")
```

## Accuracy

```{r length_stratified_accuracy}
run_length_stratified_analysis(qa_ratings_with_lengths, "accuracy") %>%
  format_length_stratified_table("accuracy")
```

## Authoritativeness

```{r length_stratified_authoritativeness}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "authoritativeness") %>%
  format_length_stratified_table("authoritativeness")
```

## Completeness

```{r length_stratified_completeness}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "completeness") %>%
  format_length_stratified_table("completeness")
```

## Verifiability

```{r length_stratified_verifiability}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "verifiability") %>%
  format_length_stratified_table("verifiability")
```


# Exploratory analysis: Plackett-Luce rankings with ties

We fit a Plackett-Luce model with a tie parameter using the
`PlackettLuce` package (Turner et al. 2020). Like the Rao-Kupper model,
this extends Bradley-Terry to handle ties, but with a different
functional form: ties are modelled by a single parameter $\delta_2$
that multiplies the geometric mean of the two vendors' worths
(Davidson 1970; Hunter 2004):

$$
P(i \succ j) = \frac{\alpha_i}{\alpha_i + \alpha_j + \delta_2 \sqrt{\alpha_i \alpha_j}}, \qquad
P(\text{tie}) = \frac{\delta_2 \sqrt{\alpha_i \alpha_j}}{\alpha_i + \alpha_j + \delta_2 \sqrt{\alpha_i \alpha_j}}.
$$

Larger $\delta_2$ implies more ties; $\delta_2 = 0$ recovers
Bradley-Terry. We bootstrap by resampling questions to get 95%
confidence intervals on the normalized worth
$\tilde{\alpha}_i = \alpha_i / \sum_k \alpha_k$ and on $\delta_2$. All
fits below are restricted to the `qa_text_only` and
`qa_text_citations` rendering modes (the `qa_full_render` condition is
excluded).

```{r pl_helpers}
library(PlackettLuce)

build_pl_rankings <- function(df, vendors) {
  K <- length(vendors)
  n <- nrow(df)
  R <- matrix(0L, nrow = n, ncol = K, dimnames = list(NULL, vendors))
  a_idx <- match(df$slot_a_vendor, vendors)
  b_idx <- match(df$slot_b_vendor, vendors)
  win_a <- !df$is_tie & df$winner == df$slot_a_vendor
  win_b <- !df$is_tie & df$winner == df$slot_b_vendor
  R[cbind(seq_len(n), a_idx)] <- ifelse(df$is_tie, 1L,
                                  ifelse(win_a, 1L, 2L))
  R[cbind(seq_len(n), b_idx)] <- ifelse(df$is_tie, 1L,
                                  ifelse(win_b, 1L, 2L))
  as.rankings(R)
}

fit_pl <- function(df, vendors = NULL) {
  if (is.null(vendors)) {
    vendors <- sort(unique(c(df$slot_a_vendor, df$slot_b_vendor)))
  }
  R   <- build_pl_rankings(df, vendors)
  fit <- PlackettLuce(R)

  cc        <- coef(fit, log = FALSE)
  tie_param <- if ("tie2" %in% names(cc)) unname(cc["tie2"]) else NA_real_
  worth     <- cc[setdiff(names(cc), "tie2")][vendors]

  list(
    vendors = vendors,
    worth   = unname(worth) / sum(worth),
    tie     = tie_param,
    fit     = fit
  )
}

bootstrap_pl <- function(df, n_boot = 500, ci_level = 0.95, seed = 42) {
  set.seed(seed)
  df_clean <- df %>% filter(!is.na(choice))
  vendors  <- sort(unique(c(df_clean$slot_a_vendor, df_clean$slot_b_vendor)))
  K        <- length(vendors)

  fit <- fit_pl(df_clean, vendors)

  q_to_rows <- split(seq_len(nrow(df_clean)),
                     as.character(df_clean$question_id))
  questions <- names(q_to_rows)

  boot_worth <- matrix(NA_real_, nrow = n_boot, ncol = K)
  boot_tie   <- rep(NA_real_, n_boot)

  for (b in seq_len(n_boot)) {
    samp_q  <- sample(questions, length(questions), replace = TRUE)
    rows    <- unlist(q_to_rows[samp_q], use.names = FALSE)
    boot_df <- df_clean[rows, ]
    if (!all(vendors %in% c(boot_df$slot_a_vendor, boot_df$slot_b_vendor))) next
    boot_fit <- tryCatch(fit_pl(boot_df, vendors), error = function(e) NULL)
    if (is.null(boot_fit)) next
    boot_worth[b, ] <- boot_fit$worth
    boot_tie[b]     <- boot_fit$tie
  }

  alpha    <- (1 - ci_level) / 2
  worth_ci <- t(apply(boot_worth, 2, quantile,
                      probs = c(alpha, 1 - alpha), na.rm = TRUE))
  tie_ci   <- quantile(boot_tie,
                       probs = c(alpha, 1 - alpha), na.rm = TRUE)

  estimates <- tibble(
    vendor   = vendors,
    worth    = fit$worth,
    ci_lower = worth_ci[, 1],
    ci_upper = worth_ci[, 2]
  ) %>%
    arrange(desc(worth)) %>%
    mutate(rank = row_number()) %>%
    select(rank, everything())

  list(
    estimates      = estimates,
    tie            = fit$tie,
    tie_ci         = tie_ci,
    n_comparisons  = nrow(df_clean),
    n_questions    = length(questions),
    n_boot         = n_boot,
    n_boot_success = sum(!is.na(boot_tie))
  )
}

format_pl_table <- function(pl_result, caption) {
  caption_full <- sprintf(
    "%s — δ₂ = %.2f [%.2f, %.2f]; N = %d comparisons across %d questions",
    caption,
    pl_result$tie,
    pl_result$tie_ci[1],
    pl_result$tie_ci[2],
    pl_result$n_comparisons,
    pl_result$n_questions
  )
  pl_result$estimates %>%
    mutate(worth = sprintf("%.3f [%.3f, %.3f]",
                           worth, ci_lower, ci_upper)) %>%
    select(rank, vendor, worth) %>%
    kable(caption   = caption_full,
          col.names = c("Rank", "Vendor",
                        "Normalized worth π̃ [95% CI]")) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Pooled across axes

```{r pl_pooled}
pl_pooled <- bootstrap_pl(qa_ratings_included)
format_pl_table(
  pl_pooled,
  "Plackett-Luce rankings, pooled across rating axes"
)
```

## Per axis

```{r pl_per_axis, results='asis'}
for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  pl_axis <- bootstrap_pl(qa_ratings_included %>% filter(axis == ax))
  print(format_pl_table(
    pl_axis,
    sprintf("Plackett-Luce rankings on axis = %s", ax)
  ))
}
```

## Cross-check: Davidson model via BradleyTerry2 + gnm

The Plackett-Luce model with ties of size two (above) is the standard
Davidson (1970) tie model. As a sanity check we refit it via
`BradleyTerry2::GenDavidson()` — a non-linear term plugged into
`gnm::gnm()` on the trinomial (win / tie / loss) expansion of the
contests — and confirm we recover the same normalized worths and tie
parameter. BradleyTerry2 parameterises the tie via `tie.max`, the
logit of the maximum tie probability; the equivalent PL δ₂ is
$\delta_2 = 2\,\text{plogis}(\text{tie.max}) / (1 - \text{plogis}(\text{tie.max}))$.

```{r davidson_helpers}
suppressMessages({
  library(BradleyTerry2)
  library(gnm)
})

fit_davidson <- function(df) {
  df      <- df %>% filter(!is.na(choice))
  vendors <- sort(unique(c(df$slot_a_vendor, df$slot_b_vendor)))
  win_a   <- !df$is_tie & df$winner == df$slot_a_vendor
  win_b   <- !df$is_tie & df$winner == df$slot_b_vendor

  expanded <- data.frame(
    match_id = factor(rep(seq_len(nrow(df)), each = 3)),
    player1  = factor(rep(df$slot_a_vendor, each = 3), levels = vendors),
    player2  = factor(rep(df$slot_b_vendor, each = 3), levels = vendors),
    result   = rep(c(1L, 0L, -1L), times = nrow(df)),
    count    = as.integer(c(t(cbind(win_a, df$is_tie, win_b))))
  )

  fit <- gnm(
    count ~ GenDavidson(result == 1, result == 0, result == -1,
                        player1, player2,
                        home.adv = NULL, tie.max = ~1) - 1,
    eliminate = match_id, family = poisson, data = expanded,
    verbose = FALSE
  )

  cc        <- coef(fit)
  tie_max   <- as.numeric(cc[grep("tie\\.max", names(cc))[1]])
  p_tie_max <- plogis(tie_max)
  delta_2   <- 2 * p_tie_max / (1 - p_tie_max)

  non_ability <- c("tie.max", "tie.scale", "tie.mode", "home.adv")
  ability_cc  <- cc[!names(cc) %in% non_ability]
  log_alpha   <- setNames(rep(NA_real_, length(vendors)), vendors)
  for (v in vendors) {
    idx <- which(endsWith(names(ability_cc), v))
    if (length(idx) > 0) log_alpha[v] <- as.numeric(ability_cc[idx[1]])
  }
  log_alpha[is.na(log_alpha)] <- 0  # gnm aliases the reference level
  alpha <- exp(log_alpha)

  list(
    vendors   = vendors,
    worth     = unname(alpha / sum(alpha)),
    delta_2   = delta_2,
    p_tie_max = p_tie_max,
    fit       = fit
  )
}
```

### Comparison: pooled across axes

```{r davidson_compare_pooled}
dav_pooled <- fit_davidson(qa_ratings_included)

tibble(
  vendor   = pl_pooled$estimates$vendor,
  pl_worth = pl_pooled$estimates$worth
) %>%
  left_join(
    tibble(vendor = dav_pooled$vendors, dav_worth = dav_pooled$worth),
    by = "vendor"
  ) %>%
  arrange(desc(pl_worth)) %>%
  mutate(
    rank      = row_number(),
    diff      = pl_worth - dav_worth,
    pl_worth  = sprintf("%.4f", pl_worth),
    dav_worth = sprintf("%.4f", dav_worth),
    diff      = sprintf("%+.4f", diff)
  ) %>%
  select(rank, vendor, pl_worth, dav_worth, diff) %>%
  kable(
    caption = sprintf(
      paste0("PlackettLuce vs. Davidson (BradleyTerry2 + gnm), ",
             "pooled across axes. δ₂(PL) = %.3f, ",
             "δ₂(Davidson) = %.3f."),
      pl_pooled$tie, dav_pooled$delta_2
    ),
    col.names = c("Rank", "Vendor",
                  "Worth (PlackettLuce)", "Worth (Davidson)",
                  "PL − Davidson")
  ) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "hover"))
```

