---
title: "Pilot Data: Descriptive Statistics & Win Rates"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(lme4)
library(broom.mixed)

source("favorability_test.R")
source("data_helpers.R")
```

# Load data

Vendor identities are blinded by `blind_vendors.R`, which writes the
`blinded_*.csv` files read below and a `secret_mapping.txt` file (kept
outside the analysis pipeline). Run
`Rscript blind_vendors.R pilot_data secret/secret_mapping.txt`
before knitting this document.

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users)
```

# Primary Analysis: Favorability Score Tests

One-vs-rest favorability score test for each vendor on the two primary
outcome axes (accuracy and clinical_utility), restricted to the text-only
rendering condition. Bootstrap resamples questions as clusters.

```{r favorability_test_helpers}
prepare_one_vs_rest_df <- function(df, vendor) {
  df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      outcome = case_when(
        winner == vendor ~ 1L,
        loser == vendor ~ -1L,
        is_tie ~ 0L
      )
    ) %>%
    select(question_id, outcome)
}

run_favorability_analysis <- function(df, axis_name,
                                      subversion_filter = "qa_text_only",
                                      n_boot = 2000, seed = 42) {
  set.seed(seed)

  axis_df <- df %>%
    filter(subversion == subversion_filter, axis == axis_name)

  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    test_df <- prepare_one_vs_rest_df(axis_df, v)

    fav_score <- get_favorability_score(test_df$outcome)
    test_result <- run_favorability_test(test_df, n_boot = n_boot, ci_level=0.95)

    tibble(
      vendor = v,
      wins = sum(test_df$outcome == 1),
      ties = sum(test_df$outcome == 0),
      losses = sum(test_df$outcome == -1),
      n = nrow(test_df),
      favorability = fav_score,
      ci_lower = test_result$ci_lower,
      ci_upper = test_result$ci_upper,
      p_value = test_result$p_value
    )
  })
}

format_favorability_table <- function(results, axis_name) {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
      p_value = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
    kable(caption = sprintf(
      "One-vs-rest favorability score test (text-only, %s axis)",
      axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility

```{r favorability_test_clinical}
run_favorability_analysis(qa_ratings, "clinical_utility") %>%
  format_favorability_table("clinical utility")
```

## Accuracy

```{r favorability_test_accuracy}
run_favorability_analysis(qa_ratings, "accuracy") %>%
  format_favorability_table("accuracy")
```

## Authoritativeness

```{r favorability_test_authoritativeness}
run_favorability_analysis(qa_ratings, "authoritativeness") %>%
  format_favorability_table("authoritativeness")
```

## Completeness

```{r favorability_test_completeness}
run_favorability_analysis(qa_ratings, "completeness") %>%
  format_favorability_table("completeness")
```

## Verifiability

```{r favorability_test_verifiability}
run_favorability_analysis(qa_ratings, "verifiability") %>%
  format_favorability_table("verifiability")
```

# Secondary analysis: Win-rate tables

For every (assignment, axis) row we know which two vendors were compared
and the chosen winner. The matrix below reports, for each ordered pair
(row vendor vs. column vendor), the row vendor's win rate across all
axes. Ties are *included* in the denominator (and count as 0 wins for
both vendors), so cell[row, col] + cell[col, row] + tie-rate = 100%.

## Win-rate matrix (aggregate)

```{r winrate_matrix}
build_win_matrix <- function(df, n_boot = 1000) {
  df %>%
    mutate(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_win = !is_tie & winner == slot_a_vendor
    ) %>%
    bind_rows(
      df %>%
        mutate(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_win = !is_tie & winner == slot_b_vendor
        )
    ) %>%
    select(question_id, row_vendor, col_vendor, row_win) %>%
    group_by(row_vendor, col_vendor) %>%
    group_modify(~ {
      res <- run_win_rate_ci(
        data.frame(question_id = .x$question_id, row_win = .x$row_win),
        n_boot = n_boot
      )
      tibble(
        n        = res$n,
        wins     = res$wins,
        win_rate = res$win_rate,
        ci_lower = res$ci_lower,
        ci_upper = res$ci_upper
      )
    }) %>%
    ungroup() %>%
    mutate(cell = sprintf("%.1f%% [%.1f, %.1f] (%d/%d)",
                          100 * win_rate,
                          100 * ci_lower,
                          100 * ci_upper,
                          wins, n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
                values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}

set.seed(42)

kable(
  build_win_matrix(qa_ratings),
  caption = paste0(
    "Row vendor's win rate vs. column vendor (ties included in ",
    "denominator). Cell = win% [95% CI] (wins / all comparisons). ",
    "CI from question-clustered bootstrap (1000 resamples)."
  )
) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Win-rate matrices per axis

Same construction as the pooled matrix, repeated within each rating axis.

```{r winrate_matrix_per_axis, results='asis'}
set.seed(42)

axes <- sort(unique(qa_ratings$axis))

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor on axis = %s ",
               "(ties included). Cell = win%% [95%% CI] (wins/N)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

## Win-rate matrices per rendering mode

Same construction as the pooled matrix, repeated within each rendering
mode (`subversion` in `users.csv`).

```{r winrate_matrix_per_subversion, results='asis'}
subversions <- sort(unique(qa_ratings$subversion))

for (sv in subversions) {
  cat(sprintf("\n\n### Rendering mode: %s\n\n", sv))
  mat <- build_win_matrix(qa_ratings %>% filter(subversion == sv))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor for ",
               "rendering mode = %s (ties included in denominator)."),
        sv
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

# Sensitivity analysis: Win-margin tables

Same row-vs-column structure as the win-rate matrices, but each cell is
the mean number of Likert units the row vendor beat the column vendor
by. The Likert scale is `strongly_*` = 2 units, `slightly_*` = 1 unit,
`tie` = 0 units. Ties are *included* (they contribute 0 units). Cells
are mirrored: `cell[row, col] = -cell[col, row]`. Cell format:
`mean (net/N)` where `net` is the signed sum of units across all
comparisons in that cell.

```{r margin_helpers}
likert_units_a <- function(choice) {
  case_when(
    choice == "strongly_a" ~  2,
    choice == "slightly_a" ~  1,
    choice == "tie"        ~  0,
    choice == "slightly_b" ~ -1,
    choice == "strongly_b" ~ -2
  )
}

build_margin_matrix <- function(df) {
  df %>%
    mutate(units_a = likert_units_a(choice)) %>%
    transmute(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_units  = units_a
    ) %>%
    bind_rows(
      df %>%
        mutate(units_a = likert_units_a(choice)) %>%
        transmute(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_units  = -units_a
        )
    ) %>%
    group_by(row_vendor, col_vendor) %>%
    summarise(
      n = n(),
      net_units  = sum(row_units),
      mean_units = net_units / n,
      .groups = "drop"
    ) %>%
    mutate(cell = sprintf("%+.2f (%+d/%d)",
                          mean_units,
                          as.integer(net_units),
                          n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
                values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}
```

## Win-margin matrix

```{r margin_matrix}
kable(
  build_margin_matrix(qa_ratings),
  caption = paste0(
    "Mean Likert-unit margin the row vendor won by vs. the column ",
    "vendor (ties included as 0). Cell = mean (net/N)."
  )
) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "hover"))
```

## Win-margin matrices per axis

```{r margin_matrix_per_axis, results='asis'}
for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_margin_matrix(qa_ratings %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Mean Likert-unit margin row vendor won by vs. column ",
               "vendor on axis = %s (ties included)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

## Win-margin matrices per rendering mode

```{r margin_matrix_per_subversion, results='asis'}
for (sv in subversions) {
  cat(sprintf("\n\n### Rendering mode: %s\n\n", sv))
  mat <- build_margin_matrix(qa_ratings %>% filter(subversion == sv))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Mean Likert-unit margin row vendor won by vs. column ",
               "vendor for rendering mode = %s (ties included)."),
        sv
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

