---
title: "Statistical Analysis: Win Rates"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
  output_dir: null
  output_name: "analysis"
  bootstrap_by: "question"
  drop_vendor: null
  user_stratification: null
  equal_weight: false
  unblind: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom.mixed)

source("favorability_test.R")
source("data_helpers.R")
```

# Analysis parameters

```{r params, echo=FALSE, results='asis'}
cat(sprintf("- **Data directory:** `%s`\n", params$data_dir))
cat(sprintf("- **Output directory:** `%s`\n", ifelse(is.null(params$output_dir), params$data_dir, params$output_dir)))
cat(sprintf("- **Bootstrap by:** %s\n", params$bootstrap_by))
cat(sprintf("- **User stratification:** %s\n", ifelse(is.null(params$user_stratification), "none", params$user_stratification)))
cat(sprintf("- **Drop vendor:** %s\n", ifelse(is.null(params$drop_vendor), "none", params$drop_vendor)))
cat(sprintf("- **Equal weight:** %s\n", ifelse(params$equal_weight, "yes (vendor pairs weighted equally)", "no")))
cat(sprintf("- **Unblind vendors:** %s\n", ifelse(params$unblind, "yes", "no")))
```

# Load data

Vendor identities are blinded by `blind_vendors.R`, which writes the
`blinded_*.csv` files read below and a `secret_mapping.csv` file (kept
outside the analysis pipeline). Run
`Rscript blind_vendors.R pilot_data secret/secret_mapping.csv`
before knitting this document.

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

# Vendor display names: blinded (A/B/C/D) by default, unblinded if params$unblind
vendor_display_names <- c("A" = "A", "B" = "B", "C" = "C", "D" = "D")
if (params$unblind) {
  vendor_mapping <- read_csv(file.path(data_dir, "secret_mapping.csv"))
  for (i in seq_len(nrow(vendor_mapping))) {
    code <- vendor_mapping$vendor[i]
    name <- vendor_mapping$provider[i]
    vendor_display_names[code] <- case_when(
      grepl("openevidence", name, ignore.case = TRUE) ~ "OE",
      grepl("gpt", name, ignore.case = TRUE) ~ "GPT",
      grepl("claude|opus", name, ignore.case = TRUE) ~ "Claude",
      grepl("gemini", name, ignore.case = TRUE) ~ "Gemini",
      TRUE ~ name
    )
  }
}

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
user_registrations  <- read_csv(file.path(data_dir, "recruitment_combined.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
users <- get_eligible_users(users, user_registrations, user_stratification=params$user_stratification)
prior_users_path <- file.path(data_dir, "blinded_prior_users.csv")
prior_users <- if (file.exists(prior_users_path)) {
  read_csv(prior_users_path)
} else {
  NULL
}
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users, prior_users = prior_users)
qa_ratings <- apply_inclusion_criteria(qa_ratings)

if (!is.null(params$drop_vendor)) {
  stopifnot(params$drop_vendor %in% c("A", "B", "C", "D"))
  qa_ratings <- qa_ratings %>%
    filter(slot_a_vendor != params$drop_vendor,
           slot_b_vendor != params$drop_vendor)
  cat(sprintf("Dropped vendor %s from analysis\n", params$drop_vendor))
}
```

# Primary Analysis: Win difference tests

One-vs-rest win difference (favorability score) test for each vendor on the two primary
outcome axes (accuracy and clinical_utility), restricted to the text-only
rendering condition. Bootstrap resamples questions as clusters.

```{r favorability_test_helpers}
compute_pair_weights <- function(df) {
  pair_counts <- df %>%
    mutate(pair = paste0(pmin(slot_a_vendor, slot_b_vendor), "_",
                         pmax(slot_a_vendor, slot_b_vendor))) %>%
    count(pair, name = "pair_n")

  df %>%
    mutate(pair = paste0(pmin(slot_a_vendor, slot_b_vendor), "_",
                         pmax(slot_a_vendor, slot_b_vendor))) %>%
    left_join(pair_counts, by = "pair") %>%
    mutate(weight = 1 / pair_n) %>%
    select(-pair, -pair_n)
}

prepare_one_vs_rest_df <- function(df, vendor, use_weights = FALSE) {
  result <- df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      outcome = case_when(
        winner == vendor ~ 1L,
        loser == vendor ~ -1L,
        is_tie ~ 0L
      )
    )

  if (use_weights) {
    result <- compute_pair_weights(result) %>%
      select(question_id, user_id, outcome, weight)
  } else {
    result <- result %>% select(question_id, user_id, outcome)
  }
  result
}

run_favorability_analysis <- function(df, axis_name,
                                      subversion_filter = "qa_text_only",
                                      n_boot = 10000, seed = 42,
                                      bootstrap_by = params$bootstrap_by,
                                      use_weights = params$equal_weight) {
  set.seed(seed)

  axis_df <- df %>%
    filter(subversion == subversion_filter, axis == axis_name)

  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    test_df <- prepare_one_vs_rest_df(axis_df, v, use_weights = use_weights)

    weights <- if (use_weights) test_df$weight else NULL
    fav_score <- get_favorability_score(test_df$outcome, weights)
    test_result <- run_favorability_test(test_df, n_boot = n_boot, ci_level=0.95,
                                         bootstrap_by = bootstrap_by)

    tibble(
      vendor = v,
      wins = sum(test_df$outcome == 1),
      ties = sum(test_df$outcome == 0),
      losses = sum(test_df$outcome == -1),
      n = nrow(test_df),
      favorability = fav_score,
      ci_lower = test_result$ci_lower,
      ci_upper = test_result$ci_upper,
      p_value = test_result$p_value
    )
  })
}

format_favorability_table <- function(results, axis_name,
                                      rendering = "text-only") {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
      p_value = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
    kable(caption = sprintf(
      "One-vs-rest favorability score test (%s, %s axis)",
      rendering, axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility

```{r favorability_test_clinical}
favorability_clinical <- run_favorability_analysis(qa_ratings, "clinical_utility")
format_favorability_table(favorability_clinical, "clinical utility")
```

## Accuracy

```{r favorability_test_accuracy}
favorability_accuracy <- run_favorability_analysis(qa_ratings, "accuracy")
format_favorability_table(favorability_accuracy, "accuracy")
```

## Authoritativeness

```{r favorability_test_authoritativeness}
favorability_authoritativeness <- run_favorability_analysis(qa_ratings, "authoritativeness")
format_favorability_table(favorability_authoritativeness, "authoritativeness")
```

## Completeness

```{r favorability_test_completeness}
favorability_completeness <- run_favorability_analysis(qa_ratings, "completeness")
format_favorability_table(favorability_completeness, "completeness")
```

## Verifiability

```{r favorability_test_verifiability}
favorability_verifiability <- run_favorability_analysis(qa_ratings, "verifiability")
format_favorability_table(favorability_verifiability, "verifiability")
```

## Summary: Outcome distribution by vendor and axis

```{r outcome_distribution_plot, fig.width=10, fig.height=6}
compute_outcome_distribution <- function(df, axis_name,
                                         subversion_filter = "qa_text_only",
                                         use_weights = params$equal_weight) {
  axis_df <- df %>%
    filter(subversion == subversion_filter, axis == axis_name)

  if (use_weights) {
    axis_df <- compute_pair_weights(axis_df)
  } else {
    axis_df$weight <- 1
  }

  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    vendor_df <- axis_df %>%
      filter(slot_a_vendor == v | slot_b_vendor == v) %>%
      mutate(
        outcome = case_when(
          slot_a_vendor == v & choice == "strongly_a" ~ "Strong Win",
          slot_b_vendor == v & choice == "strongly_b" ~ "Strong Win",
          slot_a_vendor == v & choice == "slightly_a" ~ "Slight Win",
          slot_b_vendor == v & choice == "slightly_b" ~ "Slight Win",
          choice == "tie" ~ "Tie",
          slot_a_vendor == v & choice == "slightly_b" ~ "Slight Loss",
          slot_b_vendor == v & choice == "slightly_a" ~ "Slight Loss",
          slot_a_vendor == v & choice == "strongly_b" ~ "Strong Loss",
          slot_b_vendor == v & choice == "strongly_a" ~ "Strong Loss"
        )
      )

    vendor_df %>%
      group_by(outcome) %>%
      summarise(weighted_n = sum(weight), .groups = "drop") %>%
      mutate(
        vendor = v,
        axis = axis_name,
        pct = 100 * weighted_n / sum(weighted_n)
      )
  })
}

axes <- c("accuracy", "clinical_utility", "authoritativeness", "verifiability", "completeness")

outcome_data <- map_dfr(axes, function(ax) {
  compute_outcome_distribution(qa_ratings, ax)
})

outcome_data <- outcome_data %>%
  mutate(
    outcome = factor(outcome,
                     levels = c("Strong Loss", "Slight Loss", "Tie",
                                "Slight Win", "Strong Win"))
  )

vendor_order <- outcome_data %>%
  filter(outcome %in% c("Slight Win", "Strong Win")) %>%
  group_by(vendor) %>%
  summarise(avg_win_pct = mean(pct), .groups = "drop") %>%
  arrange(desc(avg_win_pct)) %>%
  pull(vendor)

# Create display name order matching vendor_order
vendor_order_display <- vendor_display_names[vendor_order]

outcome_data <- outcome_data %>%
  mutate(vendor_display = vendor_display_names[vendor]) %>%
  mutate(vendor_display = factor(vendor_display, levels = vendor_order_display))

axis_labels <- c(
  "accuracy" = "Accuracy",
  "clinical_utility" = "Clinical Utility",
  "authoritativeness" = "Source Quality",
  "verifiability" = "Verifiability",
  "completeness" = "Completeness"
)

outcome_data <- outcome_data %>%
  mutate(axis = factor(axis_labels[axis],
                       levels = c("Accuracy", "Clinical Utility", "Source Quality",
                                  "Verifiability", "Completeness")))

sig_data <- bind_rows(
  favorability_accuracy %>% mutate(axis = "Accuracy"),
  favorability_clinical %>% mutate(axis = "Clinical Utility"),
  favorability_authoritativeness %>% mutate(axis = "Source Quality"),
  favorability_verifiability %>% mutate(axis = "Verifiability"),
  favorability_completeness %>% mutate(axis = "Completeness")
) %>%
  filter(p_value < 0.05) %>%
  mutate(
    vendor_display = vendor_display_names[vendor],
    vendor_display = factor(vendor_display, levels = vendor_order_display),
    axis = factor(axis, levels = c("Accuracy", "Clinical Utility", "Source Quality",
                                   "Verifiability", "Completeness"))
  )

p_outcome <- ggplot(outcome_data, aes(x = vendor_display, y = pct, fill = outcome)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    data = sig_data,
    aes(x = vendor_display, y = 101, label = "*"),
    inherit.aes = FALSE,
    size = 6, vjust = 0.5
  ) +
  facet_wrap(~ axis, nrow = 1, strip.position = "bottom") +
  scale_fill_manual(
    values = c("Strong Loss" = "#d73027",
               "Slight Loss" = "#fc8d59",
               "Tie" = "#fee090",
               "Slight Win" = "#91bfdb",
               "Strong Win" = "#4575b4"),
    name = "Outcome"
  ) +
  labs(
    x = "",
    y = "Outcome Distribution"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, margin = margin(t = -5)),
    strip.text.x = element_text(size = 10),
    strip.placement = "outside",
    legend.position = "bottom",
    legend.margin = margin(t = -10),
    panel.grid = element_blank(),
    plot.margin = margin(t = 0, r = 5, b = 5, l = 5)
  ) +
  coord_cartesian(ylim = c(0, 100), clip = "off")

p_outcome

ggsave(file.path(params$output_dir, "fig_outcome_distribution.pdf"),
       p_outcome, width = 10, height = 6)
```

# Secondary analysis: Win-rate tables

For every grader, answers from two vendors to the same question were compared.
The matrix below reports, for each pair (row vendor vs. column vendor),
the row vendor's win rate across all axes. Ties are *included* in the
denominator (and count as 0 wins for both vendors), so
cell[row, col] + cell[col, row] + tie-rate = 100%.

These matrices, the win-margin tables below, and the Davidson rankings
are all restricted to the `qa_text_only` and `qa_text_citations`
rendering modes.

```{r included_subversions}
included_subversions <- c("qa_text_only", "qa_text_citations")
qa_ratings_included <- qa_ratings %>%
  filter(subversion %in% included_subversions)
```

```{r winrate_matrix}
build_win_matrix <- function(df, n_boot = 10000,
                             bootstrap_by = params$bootstrap_by) {
  df %>%
    mutate(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_win = !is_tie & winner == slot_a_vendor
    ) %>%
    bind_rows(
      df %>%
        mutate(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_win = !is_tie & winner == slot_b_vendor
        )
    ) %>%
    select(question_id, user_id, row_vendor, col_vendor, row_win) %>%
    group_by(row_vendor, col_vendor) %>%
    group_modify(~ {
      res <- run_win_rate_ci(
        data.frame(question_id = .x$question_id, user_id = .x$user_id,
                   row_win = .x$row_win),
        n_boot = n_boot, bootstrap_by = bootstrap_by
      )
      tibble(
        n        = res$n,
        wins     = res$wins,
        win_rate = res$win_rate,
        ci_lower = res$ci_lower,
        ci_upper = res$ci_upper
      )
    }) %>%
    ungroup() %>%
    mutate(cell = sprintf("%.1f%% [%.1f, %.1f] (%d/%d)",
                          100 * win_rate,
                          100 * ci_lower,
                          100 * ci_upper,
                          wins, n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
                values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}

build_win_matrix_for_plot <- function(df, n_boot = 10000,
                                      bootstrap_by = params$bootstrap_by) {
  df %>%
    mutate(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_win = !is_tie & winner == slot_a_vendor
    ) %>%
    bind_rows(
      df %>%
        mutate(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_win = !is_tie & winner == slot_b_vendor
        )
    ) %>%
    select(question_id, user_id, row_vendor, col_vendor, row_win) %>%
    group_by(row_vendor, col_vendor) %>%
    group_modify(~ {
      res <- run_win_rate_ci(
        data.frame(question_id = .x$question_id, user_id = .x$user_id,
                   row_win = .x$row_win),
        n_boot = n_boot, bootstrap_by = bootstrap_by
      )
      tibble(
        n        = res$n,
        wins     = res$wins,
        win_rate = res$win_rate,
        ci_lower = res$ci_lower,
        ci_upper = res$ci_upper
      )
    }) %>%
    ungroup()
}
```

## Win-rate matrices per axis

Same construction as the pooled matrix, repeated within each rating axis.

```{r winrate_matrix_per_axis, results='asis'}
set.seed(42)

axes <- sort(unique(qa_ratings$axis))
winrate_matrices_per_axis <- list()

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings_included %>% filter(axis == ax))
  winrate_matrices_per_axis[[ax]] <- mat
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor on axis = %s ",
               "(ties included). Cell = win%% [95%% CI] (wins/N)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

## Win-rate matrix heatmaps

```{r winrate_heatmap_helper}
axis_display_names <- c(

  "accuracy" = "Accuracy",
  "clinical_utility" = "Clinical Utility",
  "authoritativeness" = "Source Quality",
  "verifiability" = "Verifiability",
  "completeness" = "Completeness"
)

plot_winrate_heatmap <- function(df, axis_name) {
  plot_data <- build_win_matrix_for_plot(df %>% filter(axis == axis_name))

  plot_data <- plot_data %>%
    mutate(
      row_display = vendor_display_names[row_vendor],
      col_display = vendor_display_names[col_vendor],
      label_rate = sprintf("%.1f%%", 100 * win_rate),
      label_ci = sprintf("[%.1f, %.1f]", 100 * ci_lower, 100 * ci_upper),
      text_color = ifelse(win_rate > 0.3, "white", "black")
    )

  display_name <- axis_display_names[axis_name]

ggplot(plot_data, aes(x = col_display, y = row_display, fill = win_rate)) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(aes(label = label_rate, fontface = "bold"), size = 4, color = "black",
              nudge_y = 0.12) +
    geom_text(aes(label = label_ci), size = 2.5, color = "black",
              nudge_y = -0.15) +
    scale_fill_gradient(
      low = "#f7fbff", high = "#2171b5",
      limits = c(0, 0.5),
      labels = scales::percent,
      name = "Win Rate",
      oob = scales::squish
    ) +
    scale_y_discrete(limits = rev) +
    labs(
      x = paste0("... vs. this system"),
      y = paste0(display_name, " win rate of this system...")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 0, hjust = 0.5),
      panel.grid = element_blank(),
      legend.position = "none"
    ) +
    coord_fixed()
}
```

### Accuracy

```{r winrate_heatmap_accuracy, fig.width=6, fig.height=5}
p_heatmap_accuracy <- plot_winrate_heatmap(qa_ratings_included, "accuracy")
p_heatmap_accuracy
ggsave(file.path(params$output_dir, "fig_winrate_heatmap_accuracy.pdf"),
       p_heatmap_accuracy, width = 6, height = 5)
```

### Clinical Utility

```{r winrate_heatmap_clinical, fig.width=6, fig.height=5}
plot_winrate_heatmap(qa_ratings_included, "clinical_utility")
```

### Source Quality

```{r winrate_heatmap_authoritativeness, fig.width=6, fig.height=5}
plot_winrate_heatmap(qa_ratings_included, "authoritativeness")
```

### Verifiability

```{r winrate_heatmap_verifiability, fig.width=6, fig.height=5}
plot_winrate_heatmap(qa_ratings_included, "verifiability")
```

### Completeness

```{r winrate_heatmap_completeness, fig.width=6, fig.height=5}
plot_winrate_heatmap(qa_ratings_included, "completeness")
```

# Secondary Analysis: Win difference score for `qa_text_citations`

Same one-vs-rest win difference (favorability score) test as the primary analysis, but
restricted to the `qa_text_citations` rendering mode (the primary
analysis covers `qa_text_only`).

## Clinical utility

```{r favorability_test_citations_clinical}
favorability_citations_clinical <- run_favorability_analysis(
  qa_ratings, "clinical_utility", subversion_filter = "qa_text_citations")
format_favorability_table(favorability_citations_clinical, "clinical utility",
                          rendering = "qa_text_citations")
```

## Accuracy

```{r favorability_test_citations_accuracy}
favorability_citations_accuracy <- run_favorability_analysis(
  qa_ratings, "accuracy", subversion_filter = "qa_text_citations")
format_favorability_table(favorability_citations_accuracy, "accuracy",
                          rendering = "qa_text_citations")
```

## Authoritativeness

```{r favorability_test_citations_authoritativeness}
favorability_citations_authoritativeness <- run_favorability_analysis(
  qa_ratings, "authoritativeness", subversion_filter = "qa_text_citations")
format_favorability_table(favorability_citations_authoritativeness, "authoritativeness",
                          rendering = "qa_text_citations")
```

## Completeness

```{r favorability_test_citations_completeness}
favorability_citations_completeness <- run_favorability_analysis(
  qa_ratings, "completeness", subversion_filter = "qa_text_citations")
format_favorability_table(favorability_citations_completeness, "completeness",
                          rendering = "qa_text_citations")
```

## Verifiability

```{r favorability_test_citations_verifiability}
favorability_citations_verifiability <- run_favorability_analysis(
  qa_ratings, "verifiability", subversion_filter = "qa_text_citations")
format_favorability_table(favorability_citations_verifiability, "verifiability",
                          rendering = "qa_text_citations")
```

# Sensitivity analysis: Strict win-rate tables

Same construction as the win-rate matrices above, but with a stricter
definition: a "win" requires `strongly_a`/`strongly_b` for the row
vendor, and a "loss" requires the strong preference for the column
vendor. `slightly_*` and `tie` outcomes are treated as non-wins for
both vendors. The denominator is still every comparison, so
`cell[row, col] + cell[col, row] + strict-non-win-rate = 100%`.

```{r strict_winrate_helpers}
apply_strict_win_definition <- function(df) {
  df %>%
    mutate(
      winner = case_when(
        choice == "strongly_a" ~ slot_a_vendor,
        choice == "strongly_b" ~ slot_b_vendor,
        TRUE ~ NA_character_
      ),
      loser = case_when(
        choice == "strongly_a" ~ slot_b_vendor,
        choice == "strongly_b" ~ slot_a_vendor,
        TRUE ~ NA_character_
      ),
      is_tie = !choice %in% c("strongly_a", "strongly_b")
    )
}

qa_ratings_strict <- apply_strict_win_definition(qa_ratings_included)
```

## Strict win-rate matrices per axis

```{r strict_winrate_matrix_per_axis, results='asis'}
set.seed(42)

strict_winrate_matrices_per_axis <- list()

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings_strict %>% filter(axis == ax))
  strict_winrate_matrices_per_axis[[ax]] <- mat
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's strict win rate vs. column vendor on ",
               "axis = %s (only `strongly_*` counts as a win; ",
               "denominator = all comparisons)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

# Sensitivity analysis: Length-stratified favorability tests

Re-runs the one-vs-rest favorability score test, but stratified by
whether the vendor under test produced a *longer* or *shorter* answer
than its comparator (character length from `blinded_qa_answers.csv`).
Unlike the primary analysis, rendering conditions are **pooled** here
(all `subversion` values included). The comparison count drops because
each comparison goes into exactly one stratum; comparisons where the two
vendors produced exactly equal-length answers are dropped.

```{r length_stratified_helpers}
answer_lengths <- read_csv(file.path(data_dir, "blinded_qa_answers.csv")) %>%
  select(question_id, vendor, answer_len)

qa_ratings_with_lengths <- qa_ratings %>%
  left_join(
    answer_lengths %>%
      rename(slot_a_vendor = vendor, slot_a_answer_len = answer_len),
    by = c("question_id", "slot_a_vendor")
  ) %>%
  left_join(
    answer_lengths %>%
      rename(slot_b_vendor = vendor, slot_b_answer_len = answer_len),
    by = c("question_id", "slot_b_vendor")
  )

equal_length_pairs <- qa_ratings_with_lengths %>%
  filter(slot_a_answer_len == slot_b_answer_len) %>%
  distinct(question_id, slot_a_vendor, slot_b_vendor)

n_questions_equal_length <- n_distinct(equal_length_pairs$question_id)

writeLines(sprintf(
  "Equal-length vendor pairs: %d (across %d unique questions).",
  nrow(equal_length_pairs), n_questions_equal_length
), con = stderr())
```

```{r}
prepare_one_vs_rest_df_with_lengths <- function(df, vendor, use_weights = FALSE) {
  result <- df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      vendor_len     = if_else(slot_a_vendor == vendor,
                               slot_a_answer_len, slot_b_answer_len),
      comparator_len = if_else(slot_a_vendor == vendor,
                               slot_b_answer_len, slot_a_answer_len),
      outcome = case_when(
        winner == vendor ~ 1L,
        loser  == vendor ~ -1L,
        is_tie ~ 0L
      )
    )

  if (use_weights) {
    result <- compute_pair_weights(result) %>%
      select(question_id, user_id, outcome, vendor_len, comparator_len, weight)
  } else {
    result <- result %>%
      select(question_id, user_id, outcome, vendor_len, comparator_len)
  }
  result
}

run_length_stratified_analysis <- function(df, axis_name,
                                           n_boot = 10000, seed = 42,
                                           bootstrap_by = params$bootstrap_by,
                                           use_weights = params$equal_weight) {
  set.seed(seed)

  axis_df <- df %>% filter(axis == axis_name)
  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    base_df <- prepare_one_vs_rest_df_with_lengths(axis_df, v, use_weights = use_weights) %>%
      filter(vendor_len != comparator_len)

    map_dfr(c("longer", "shorter"), function(stratum) {
      stratum_df <- base_df %>%
        filter(if (stratum == "longer") vendor_len > comparator_len
                                  else vendor_len < comparator_len)

      if (use_weights) {
        stratum_df <- stratum_df %>% select(question_id, user_id, outcome, weight)
      } else {
        stratum_df <- stratum_df %>% select(question_id, user_id, outcome)
      }

      weights <- if (use_weights) stratum_df$weight else NULL
      fav_score   <- get_favorability_score(stratum_df$outcome, weights)
      test_result <- run_favorability_test(stratum_df,
                                           n_boot = n_boot, ci_level = 0.95,
                                           bootstrap_by = bootstrap_by)

      tibble(
        vendor       = v,
        stratum      = stratum,
        wins         = sum(stratum_df$outcome == 1),
        ties         = sum(stratum_df$outcome == 0),
        losses       = sum(stratum_df$outcome == -1),
        n            = nrow(stratum_df),
        favorability = fav_score,
        ci_lower     = test_result$ci_lower,
        ci_upper     = test_result$ci_upper,
        p_value      = test_result$p_value
      )
    })
  })
}

format_length_stratified_table <- function(results, axis_name) {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95        = sprintf("[%.1f%%, %.1f%%]",
                             100 * ci_lower, 100 * ci_upper),
      p_value      = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, stratum, wins, ties, losses, n,
           favorability, ci_95, p_value) %>%
    arrange(vendor, stratum) %>%
    kable(caption = sprintf(
      paste0("Length-stratified one-vs-rest favorability test ",
             "(pooled rendering, %s axis). Stratum = vendor's answer ",
             "longer/shorter than comparator."),
      axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility

```{r length_stratified_clinical}
length_stratified_clinical <- run_length_stratified_analysis(
  qa_ratings_with_lengths, "clinical_utility")
format_length_stratified_table(length_stratified_clinical, "clinical utility")
```

## Accuracy

```{r length_stratified_accuracy}
length_stratified_accuracy <- run_length_stratified_analysis(
  qa_ratings_with_lengths, "accuracy")
format_length_stratified_table(length_stratified_accuracy, "accuracy")
```

## Authoritativeness

```{r length_stratified_authoritativeness}
length_stratified_authoritativeness <- run_length_stratified_analysis(
  qa_ratings_with_lengths, "authoritativeness")
format_length_stratified_table(length_stratified_authoritativeness, "authoritativeness")
```

## Completeness

```{r length_stratified_completeness}
length_stratified_completeness <- run_length_stratified_analysis(
  qa_ratings_with_lengths, "completeness")
format_length_stratified_table(length_stratified_completeness, "completeness")
```

## Verifiability

```{r length_stratified_verifiability}
length_stratified_verifiability <- run_length_stratified_analysis(
  qa_ratings_with_lengths, "verifiability")
format_length_stratified_table(length_stratified_verifiability, "verifiability")
```

## Summary: Length-stratified favorability scores

```{r length_stratified_favorability_data}
length_stratified_all <- bind_rows(
  length_stratified_accuracy %>% mutate(axis = "Accuracy"),
  length_stratified_clinical %>% mutate(axis = "Clinical Utility"),
  length_stratified_authoritativeness %>% mutate(axis = "Source Quality"),
  length_stratified_verifiability %>% mutate(axis = "Verifiability"),
  length_stratified_completeness %>% mutate(axis = "Completeness")
) %>%
  mutate(
    vendor = factor(vendor, levels = vendor_order),
    stratum = factor(stratum, levels = c("longer", "shorter"),
                     labels = c("Longer", "Shorter"))
  )
```

Plots are combined with OE status stratification below.

# Sensitivity analysis: OE status-stratified favorability tests

Re-runs the one-vs-rest favorability score test, stratified by whether
the rater is an active OpenEvidence user ("On OE") or not ("Off OE").
This tests whether familiarity with AI-generated medical content affects
vendor preferences.

```{r oe_status_stratified_helpers}
qa_ratings_with_oe_status <- qa_ratings %>%
  left_join(
    user_registrations %>%
      select(email = Email, oe_status = `OE Status`),
    by = c("user_id" = "email")
  ) %>%

  left_join(
    users %>% select(id, email),
    by = c("user_id" = "id")
  ) %>%
  left_join(
    user_registrations %>%
      select(email = Email, oe_status2 = `OE Status`),
    by = "email"
  ) %>%
  mutate(oe_status = coalesce(oe_status, oe_status2)) %>%
  select(-oe_status2, -email) %>%
  filter(!is.na(oe_status))

n_on_oe <- qa_ratings_with_oe_status %>%
  filter(oe_status == "On OE") %>%
  distinct(user_id) %>%
  nrow()

n_off_oe <- qa_ratings_with_oe_status %>%
  filter(oe_status == "Off OE") %>%
  distinct(user_id) %>%

  nrow()

writeLines(sprintf(
 "OE status breakdown: %d 'On OE' users, %d 'Off OE' users.",
  n_on_oe, n_off_oe
), con = stderr())
```

```{r oe_status_stratified_analysis}
run_oe_status_stratified_analysis <- function(df, axis_name,
                                              n_boot = 10000, seed = 42,
                                              bootstrap_by = params$bootstrap_by,
                                              use_weights = params$equal_weight) {
  set.seed(seed)

  axis_df <- df %>% filter(axis == axis_name)
  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    base_df <- axis_df %>%
      filter(slot_a_vendor == v | slot_b_vendor == v) %>%
      mutate(
        outcome = case_when(
          winner == v ~ 1L,
          loser == v ~ -1L,
          is_tie ~ 0L
        )
      )

    if (use_weights) {
      base_df <- compute_pair_weights(base_df)
    }

    map_dfr(c("On OE", "Off OE"), function(stratum) {
      stratum_df <- base_df %>% filter(oe_status == stratum)

      if (use_weights) {
        stratum_df <- stratum_df %>% select(question_id, user_id, outcome, weight)
      } else {
        stratum_df <- stratum_df %>% select(question_id, user_id, outcome)
      }

      weights <- if (use_weights) stratum_df$weight else NULL
      fav_score <- get_favorability_score(stratum_df$outcome, weights)
      test_result <- run_favorability_test(stratum_df,
                                           n_boot = n_boot, ci_level = 0.95,
                                           bootstrap_by = bootstrap_by)

      tibble(
        vendor       = v,
        stratum      = stratum,
        wins         = sum(stratum_df$outcome == 1),
        ties         = sum(stratum_df$outcome == 0),
        losses       = sum(stratum_df$outcome == -1),
        n            = nrow(stratum_df),
        favorability = fav_score,
        ci_lower     = test_result$ci_lower,
        ci_upper     = test_result$ci_upper,
        p_value      = test_result$p_value
      )
    })
  })
}

format_oe_status_stratified_table <- function(results, axis_name) {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95        = sprintf("[%.1f%%, %.1f%%]",
                             100 * ci_lower, 100 * ci_upper),
      p_value      = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, stratum, wins, ties, losses, n,
           favorability, ci_95, p_value) %>%
    arrange(vendor, stratum) %>%
    kable(caption = sprintf(
      paste0("OE status-stratified one-vs-rest favorability test ",
             "(pooled rendering, %s axis)."),
      axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Accuracy

```{r oe_status_stratified_accuracy}
oe_status_stratified_accuracy <- run_oe_status_stratified_analysis(
  qa_ratings_with_oe_status, "accuracy")
format_oe_status_stratified_table(oe_status_stratified_accuracy, "Accuracy")
```

## Clinical Utility

```{r oe_status_stratified_clinical}
oe_status_stratified_clinical <- run_oe_status_stratified_analysis(
  qa_ratings_with_oe_status, "clinical_utility")
format_oe_status_stratified_table(oe_status_stratified_clinical, "Clinical Utility")
```

## Source Quality

```{r oe_status_stratified_authoritativeness}
oe_status_stratified_authoritativeness <- run_oe_status_stratified_analysis(
  qa_ratings_with_oe_status, "authoritativeness")
format_oe_status_stratified_table(oe_status_stratified_authoritativeness, "Source Quality")
```

## Verifiability

```{r oe_status_stratified_verifiability}
oe_status_stratified_verifiability <- run_oe_status_stratified_analysis(
  qa_ratings_with_oe_status, "verifiability")
format_oe_status_stratified_table(oe_status_stratified_verifiability, "Verifiability")
```

## Completeness

```{r oe_status_stratified_completeness}
oe_status_stratified_completeness <- run_oe_status_stratified_analysis(
  qa_ratings_with_oe_status, "completeness")
format_oe_status_stratified_table(oe_status_stratified_completeness, "Completeness")
```

## Summary: Sensitivity analysis plots (exported individually)

```{r sensitivity_plot_helpers}
make_length_plot <- function(data, axis_name) {
  plot_data <- data %>%
    mutate(
      vendor_display = vendor_display_names[vendor],
      vendor_display = factor(vendor_display, levels = vendor_order_display),
      stratum = factor(stratum, levels = c("longer", "shorter"),
                       labels = c("Longer", "Shorter"))
    )

  ggplot(plot_data, aes(x = vendor_display, y = 100 * favorability, fill = stratum)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
    geom_errorbar(
      aes(ymin = 100 * ci_lower, ymax = 100 * ci_upper),
      position = position_dodge(width = 0.8),
      width = 0.25
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    scale_fill_manual(
      values = c("Longer" = "#7b3294", "Shorter" = "#c2a5cf"),
      name = "Answer Length"
    ) +
    labs(
      x = "System",
      y = "Win Difference (%)",
      title = sprintf("%s by Answer Length", axis_name)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 0, hjust = 0.5),
      legend.position = "bottom"
    )
}

make_oe_status_plot <- function(data, axis_name) {
  plot_data <- data %>%
    mutate(
      vendor_display = vendor_display_names[vendor],
      vendor_display = factor(vendor_display, levels = vendor_order_display),
      stratum = factor(stratum, levels = c("On OE", "Off OE"))
    )

  ggplot(plot_data, aes(x = vendor_display, y = 100 * favorability, fill = stratum)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
    geom_errorbar(
      aes(ymin = 100 * ci_lower, ymax = 100 * ci_upper),
      position = position_dodge(width = 0.8),
      width = 0.25
    ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    scale_fill_manual(
      values = c("On OE" = "#d95f02", "Off OE" = "#fee0b6"),
      name = "OE Status"
    ) +
    labs(
      x = "System",
      y = "Win Difference (%)",
      title = sprintf("%s by OE Status", axis_name)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 0, hjust = 0.5),
      legend.position = "bottom"
    )
}
```

### Accuracy sensitivity plots

```{r sensitivity_accuracy, fig.width=6, fig.height=4}
p_length_accuracy <- make_length_plot(length_stratified_accuracy, "Accuracy")
p_length_accuracy
ggsave(file.path(params$output_dir, "fig_sensitivity_length_accuracy.pdf"),
       p_length_accuracy, width = 6, height = 4)

p_oe_accuracy <- make_oe_status_plot(oe_status_stratified_accuracy, "Accuracy")
p_oe_accuracy
ggsave(file.path(params$output_dir, "fig_sensitivity_oe_accuracy.pdf"),
       p_oe_accuracy, width = 6, height = 4)
```

# Exploratory analysis: Davidson rankings with ties

We fit a Davidson (1970) tie model via `BradleyTerry2::GenDavidson()` —
a non-linear term plugged into `gnm::gnm()` on the trinomial (win /
tie / loss) expansion of the contests. Like the Rao-Kupper model, this
extends Bradley-Terry to handle ties; ties are modelled by a single
parameter $\delta_2$ that multiplies the geometric mean of the two
vendors' worths:

$$
P(i \succ j) = \frac{\alpha_i}{\alpha_i + \alpha_j + \delta_2 \sqrt{\alpha_i \alpha_j}}, \qquad
P(\text{tie}) = \frac{\delta_2 \sqrt{\alpha_i \alpha_j}}{\alpha_i + \alpha_j + \delta_2 \sqrt{\alpha_i \alpha_j}}.
$$

Larger $\delta_2$ implies more ties; $\delta_2 = 0$ recovers
Bradley-Terry. BradleyTerry2 parameterises the tie via `tie.max`, the
logit of the maximum tie probability; the equivalent
$\delta_2 = 2\,\text{plogis}(\text{tie.max}) / (1 - \text{plogis}(\text{tie.max}))$.
Individual ability coefficients are not identifiable in the
non-linear term, so we use `gnm::getContrasts()` to obtain estimable
mean-centered log abilities $\log \alpha_i - \overline{\log \alpha}$
with valid SEs. We then report
$\tilde{\alpha}_i = \exp(\log \alpha_i - \overline{\log \alpha})$,
i.e. each vendor's worth as a ratio to the geometric-mean worth, with
Wald CIs $\exp(\widehat{\log \alpha}_i \pm 1.96\,\widehat{\mathrm{SE}}_i)$.
$\tilde{\alpha} > 1$ means above-average, $< 1$ means below-average.
All fits below are restricted to the `qa_text_only` and
`qa_text_citations` rendering modes (the `qa_full_render` condition is
excluded).

```{r davidson_helpers}
suppressMessages({
  library(BradleyTerry2)
  library(gnm)
})

fit_davidson <- function(df, use_weights = params$equal_weight) {
  df      <- df %>% filter(!is.na(choice))

  if (use_weights) {
    df <- compute_pair_weights(df)
  } else {
    df$weight <- 1
  }

  vendors <- sort(unique(c(df$slot_a_vendor, df$slot_b_vendor)))
  win_a   <- !df$is_tie & df$winner == df$slot_a_vendor
  win_b   <- !df$is_tie & df$winner == df$slot_b_vendor

  expanded <- data.frame(
    match_id = factor(rep(seq_len(nrow(df)), each = 3)),
    player1  = factor(rep(df$slot_a_vendor, each = 3), levels = vendors),
    player2  = factor(rep(df$slot_b_vendor, each = 3), levels = vendors),
    result   = rep(c(1L, 0L, -1L), times = nrow(df)),
    count    = as.integer(c(t(cbind(win_a, df$is_tie, win_b)))),
    weight   = rep(df$weight, each = 3)
  )

  fit <- gnm(
    count ~ GenDavidson(result == 1, result == 0, result == -1,
                        player1, player2,
                        home.adv = NULL, tie.max = ~1) - 1,
    eliminate = match_id, family = poisson, data = expanded,
    weights = weight,
    verbose = FALSE
  )

  cc          <- coef(fit)
  tie_idx     <- grep("tie\\.|home\\.adv", names(cc))
  ability_idx <- setdiff(seq_along(cc), tie_idx)

  tie_max   <- as.numeric(cc[tie_idx[1]])
  p_tie_max <- plogis(tie_max)
  delta_2   <- 2 * p_tie_max / (1 - p_tie_max)

  # gnm does not report SEs for individual ability coefficients in
  # non-linear terms (only contrasts are identified). getContrasts gives
  # each log α centered at the mean log α, with proper SEs.
  qv <- getContrasts(fit, ability_idx)$qvframe

  log_alpha <- setNames(rep(NA_real_, length(vendors)), vendors)
  log_se    <- setNames(rep(NA_real_, length(vendors)), vendors)
  for (v in vendors) {
    idx <- which(endsWith(rownames(qv), v))
    if (length(idx) > 0) {
      log_alpha[v] <- qv$estimate[idx[1]]
      log_se[v]    <- qv$SE[idx[1]]
    }
  }

  z <- qnorm(0.975)

  list(
    vendors       = vendors,
    worth         = unname(exp(log_alpha)),
    worth_lower   = unname(exp(log_alpha - z * log_se)),
    worth_upper   = unname(exp(log_alpha + z * log_se)),
    delta_2       = delta_2,
    n_comparisons = nrow(df),
    n_questions   = length(unique(df$question_id)),
    fit           = fit
  )
}

format_davidson_table <- function(dav, caption) {
  caption_full <- sprintf(
    "%s — δ₂ = %.2f; N = %d comparisons across %d questions",
    caption, dav$delta_2, dav$n_comparisons, dav$n_questions
  )
  tibble(
    vendor   = dav$vendors,
    worth    = dav$worth,
    ci_lower = dav$worth_lower,
    ci_upper = dav$worth_upper
  ) %>%
    arrange(desc(worth)) %>%
    mutate(
      rank  = row_number(),
      worth = sprintf("%.3f [%.3f, %.3f]", worth, ci_lower, ci_upper)
    ) %>%
    select(rank, vendor, worth) %>%
    kable(caption   = caption_full,
          col.names = c("Rank", "Vendor",
                        "Worth ratio α̃ [95% CI]")) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```


```{r davidson_per_axis, results='asis'}
davidson_fits_per_axis <- list()

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  dav_axis <- fit_davidson(qa_ratings_included %>% filter(axis == ax))
  davidson_fits_per_axis[[ax]] <- dav_axis
  print(format_davidson_table(
    dav_axis,
    sprintf("Davidson rankings on axis = %s", ax)
  ))
}
```

# Save everything
```{r save_outputs}
# Save workspace for R users
rdata_path <- file.path(params$output_dir, paste0(params$output_name, ".RData"))
save.image(file = rdata_path)
message("Saved RData to: ", rdata_path)
```