---
title: "Statistical Analysis: Win Rates"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
  output_dir: null
  output_name: "analysis"
  bootstrap_by: "question"
  drop_vendor: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom.mixed)

source("favorability_test.R")
source("data_helpers.R")
```

# Analysis parameters

```{r params, echo=FALSE, results='asis'}
cat(sprintf("- **Data directory:** `%s`\n", params$data_dir))
cat(sprintf("- **Output directory:** `%s`\n", ifelse(is.null(params$output_dir), params$data_dir, params$output_dir)))
cat(sprintf("- **Bootstrap by:** %s\n", params$bootstrap_by))
cat(sprintf("- **Drop vendor:** %s\n", ifelse(is.null(params$drop_vendor), "none", params$drop_vendor)))
```

# Load data

Vendor identities are blinded by `blind_vendors.R`, which writes the
`blinded_*.csv` files read below and a `secret_mapping.csv` file (kept
outside the analysis pipeline). Run
`Rscript blind_vendors.R pilot_data secret/secret_mapping.csv`
before knitting this document.

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
users <- get_eligible_users(users)
prior_users_path <- file.path(data_dir, "blinded_prior_users.csv")
prior_users <- if (file.exists(prior_users_path)) {
  read_csv(prior_users_path)
} else {
  NULL
}
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users, prior_users = prior_users)
qa_ratings <- apply_inclusion_criteria(qa_ratings)

if (!is.null(params$drop_vendor)) {
  stopifnot(params$drop_vendor %in% c("A", "B", "C", "D"))
  qa_ratings <- qa_ratings %>%
    filter(slot_a_vendor != params$drop_vendor,
           slot_b_vendor != params$drop_vendor)
  cat(sprintf("Dropped vendor %s from analysis\n", params$drop_vendor))
}
```

# Primary Analysis: Win difference tests

One-vs-rest win difference (favorability score) test for each vendor on the two primary
outcome axes (accuracy and clinical_utility), restricted to the text-only
rendering condition. Bootstrap resamples questions as clusters.

```{r favorability_test_helpers}
prepare_one_vs_rest_df <- function(df, vendor) {
  df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      outcome = case_when(
        winner == vendor ~ 1L,
        loser == vendor ~ -1L,
        is_tie ~ 0L
      )
    ) %>%
    select(question_id, user_id, outcome)
}

run_favorability_analysis <- function(df, axis_name,
                                      subversion_filter = "qa_text_only",
                                      n_boot = 10000, seed = 42,
                                      bootstrap_by = params$bootstrap_by) {
  set.seed(seed)

  axis_df <- df %>%
    filter(subversion == subversion_filter, axis == axis_name)

  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    test_df <- prepare_one_vs_rest_df(axis_df, v)

    fav_score <- get_favorability_score(test_df$outcome)
    test_result <- run_favorability_test(test_df, n_boot = n_boot, ci_level=0.95,
                                         bootstrap_by = bootstrap_by)

    tibble(
      vendor = v,
      wins = sum(test_df$outcome == 1),
      ties = sum(test_df$outcome == 0),
      losses = sum(test_df$outcome == -1),
      n = nrow(test_df),
      favorability = fav_score,
      ci_lower = test_result$ci_lower,
      ci_upper = test_result$ci_upper,
      p_value = test_result$p_value
    )
  })
}

format_favorability_table <- function(results, axis_name,
                                      rendering = "text-only") {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
      p_value = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
    kable(caption = sprintf(
      "One-vs-rest favorability score test (%s, %s axis)",
      rendering, axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility

```{r favorability_test_clinical}
run_favorability_analysis(qa_ratings, "clinical_utility") %>%
  format_favorability_table("clinical utility")
```

## Accuracy

```{r favorability_test_accuracy}
run_favorability_analysis(qa_ratings, "accuracy") %>%
  format_favorability_table("accuracy")
```

## Authoritativeness

```{r favorability_test_authoritativeness}
run_favorability_analysis(qa_ratings, "authoritativeness") %>%
  format_favorability_table("authoritativeness")
```

## Completeness

```{r favorability_test_completeness}
run_favorability_analysis(qa_ratings, "completeness") %>%
  format_favorability_table("completeness")
```

## Verifiability

```{r favorability_test_verifiability}
run_favorability_analysis(qa_ratings, "verifiability") %>%
  format_favorability_table("verifiability")
```

# Secondary analysis: Win-rate tables

For every grader, answers from two vendors to the same question were compared.
The matrix below reports, for each pair (row vendor vs. column vendor),
the row vendor's win rate across all axes. Ties are *included* in the
denominator (and count as 0 wins for both vendors), so
cell[row, col] + cell[col, row] + tie-rate = 100%.

These matrices, the win-margin tables below, and the Davidson rankings
are all restricted to the `qa_text_only` and `qa_text_citations`
rendering modes.

```{r included_subversions}
included_subversions <- c("qa_text_only", "qa_text_citations")
qa_ratings_included <- qa_ratings %>%
  filter(subversion %in% included_subversions)
```

```{r winrate_matrix}
build_win_matrix <- function(df, n_boot = 10000,
                             bootstrap_by = params$bootstrap_by) {
  df %>%
    mutate(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_win = !is_tie & winner == slot_a_vendor
    ) %>%
    bind_rows(
      df %>%
        mutate(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_win = !is_tie & winner == slot_b_vendor
        )
    ) %>%
    select(question_id, user_id, row_vendor, col_vendor, row_win) %>%
    group_by(row_vendor, col_vendor) %>%
    group_modify(~ {
      res <- run_win_rate_ci(
        data.frame(question_id = .x$question_id, user_id = .x$user_id,
                   row_win = .x$row_win),
        n_boot = n_boot, bootstrap_by = bootstrap_by
      )
      tibble(
        n        = res$n,
        wins     = res$wins,
        win_rate = res$win_rate,
        ci_lower = res$ci_lower,
        ci_upper = res$ci_upper
      )
    }) %>%
    ungroup() %>%
    mutate(cell = sprintf("%.1f%% [%.1f, %.1f] (%d/%d)",
                          100 * win_rate,
                          100 * ci_lower,
                          100 * ci_upper,
                          wins, n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
                values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}
```

## Win-rate matrices per axis

Same construction as the pooled matrix, repeated within each rating axis.

```{r winrate_matrix_per_axis, results='asis'}
set.seed(42)

axes <- sort(unique(qa_ratings$axis))

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings_included %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor on axis = %s ",
               "(ties included). Cell = win%% [95%% CI] (wins/N)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

# Secondary Analysis: Win difference score for `qa_text_citations`

Same one-vs-rest win difference (favorability score) test as the primary analysis, but
restricted to the `qa_text_citations` rendering mode (the primary
analysis covers `qa_text_only`).

## Clinical utility

```{r favorability_test_citations_clinical}
run_favorability_analysis(qa_ratings, "clinical_utility",
                          subversion_filter = "qa_text_citations") %>%
  format_favorability_table("clinical utility",
                            rendering = "qa_text_citations")
```

## Accuracy

```{r favorability_test_citations_accuracy}
run_favorability_analysis(qa_ratings, "accuracy",
                          subversion_filter = "qa_text_citations") %>%
  format_favorability_table("accuracy",
                            rendering = "qa_text_citations")
```

## Authoritativeness

```{r favorability_test_citations_authoritativeness}
run_favorability_analysis(qa_ratings, "authoritativeness",
                          subversion_filter = "qa_text_citations") %>%
  format_favorability_table("authoritativeness",
                            rendering = "qa_text_citations")
```

## Completeness

```{r favorability_test_citations_completeness}
run_favorability_analysis(qa_ratings, "completeness",
                          subversion_filter = "qa_text_citations") %>%
  format_favorability_table("completeness",
                            rendering = "qa_text_citations")
```

## Verifiability

```{r favorability_test_citations_verifiability}
run_favorability_analysis(qa_ratings, "verifiability",
                          subversion_filter = "qa_text_citations") %>%
  format_favorability_table("verifiability",
                            rendering = "qa_text_citations")
```

# Sensitivity analysis: Strict win-rate tables

Same construction as the win-rate matrices above, but with a stricter
definition: a "win" requires `strongly_a`/`strongly_b` for the row
vendor, and a "loss" requires the strong preference for the column
vendor. `slightly_*` and `tie` outcomes are treated as non-wins for
both vendors. The denominator is still every comparison, so
`cell[row, col] + cell[col, row] + strict-non-win-rate = 100%`.

```{r strict_winrate_helpers}
apply_strict_win_definition <- function(df) {
  df %>%
    mutate(
      winner = case_when(
        choice == "strongly_a" ~ slot_a_vendor,
        choice == "strongly_b" ~ slot_b_vendor,
        TRUE ~ NA_character_
      ),
      loser = case_when(
        choice == "strongly_a" ~ slot_b_vendor,
        choice == "strongly_b" ~ slot_a_vendor,
        TRUE ~ NA_character_
      ),
      is_tie = !choice %in% c("strongly_a", "strongly_b")
    )
}

qa_ratings_strict <- apply_strict_win_definition(qa_ratings_included)
```

## Strict win-rate matrices per axis

```{r strict_winrate_matrix_per_axis, results='asis'}
set.seed(42)

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings_strict %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's strict win rate vs. column vendor on ",
               "axis = %s (only `strongly_*` counts as a win; ",
               "denominator = all comparisons)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

# Sensitivity analysis: Length-stratified favorability tests

Re-runs the one-vs-rest favorability score test, but stratified by
whether the vendor under test produced a *longer* or *shorter* answer
than its comparator (character length from `blinded_qa_answers.csv`).
Unlike the primary analysis, rendering conditions are **pooled** here
(all `subversion` values included). The comparison count drops because
each comparison goes into exactly one stratum; comparisons where the two
vendors produced exactly equal-length answers are dropped.

```{r length_stratified_helpers}
answer_lengths <- read_csv(file.path(data_dir, "blinded_qa_answers.csv")) %>%
  select(question_id, vendor, answer_len)

qa_ratings_with_lengths <- qa_ratings %>%
  left_join(
    answer_lengths %>%
      rename(slot_a_vendor = vendor, slot_a_answer_len = answer_len),
    by = c("question_id", "slot_a_vendor")
  ) %>%
  left_join(
    answer_lengths %>%
      rename(slot_b_vendor = vendor, slot_b_answer_len = answer_len),
    by = c("question_id", "slot_b_vendor")
  )

equal_length_pairs <- qa_ratings_with_lengths %>%
  filter(slot_a_answer_len == slot_b_answer_len) %>%
  distinct(question_id, slot_a_vendor, slot_b_vendor)

n_questions_equal_length <- n_distinct(equal_length_pairs$question_id)

writeLines(sprintf(
  "Equal-length vendor pairs: %d (across %d unique questions).",
  nrow(equal_length_pairs), n_questions_equal_length
), con = stderr())
```

```{r}
prepare_one_vs_rest_df_with_lengths <- function(df, vendor) {
  df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      vendor_len     = if_else(slot_a_vendor == vendor,
                               slot_a_answer_len, slot_b_answer_len),
      comparator_len = if_else(slot_a_vendor == vendor,
                               slot_b_answer_len, slot_a_answer_len),
      outcome = case_when(
        winner == vendor ~ 1L,
        loser  == vendor ~ -1L,
        is_tie ~ 0L
      )
    ) %>%
    select(question_id, user_id, outcome, vendor_len, comparator_len)
}

run_length_stratified_analysis <- function(df, axis_name,
                                           n_boot = 10000, seed = 42,
                                           bootstrap_by = params$bootstrap_by) {
  set.seed(seed)

  axis_df <- df %>% filter(axis == axis_name)
  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    base_df <- prepare_one_vs_rest_df_with_lengths(axis_df, v) %>%
      filter(vendor_len != comparator_len)

    map_dfr(c("longer", "shorter"), function(stratum) {
      stratum_df <- base_df %>%
        filter(if (stratum == "longer") vendor_len > comparator_len
                                  else vendor_len < comparator_len) %>%
        select(question_id, user_id, outcome)

      fav_score   <- get_favorability_score(stratum_df$outcome)
      test_result <- run_favorability_test(stratum_df,
                                           n_boot = n_boot, ci_level = 0.95,
                                           bootstrap_by = bootstrap_by)

      tibble(
        vendor       = v,
        stratum      = stratum,
        wins         = sum(stratum_df$outcome == 1),
        ties         = sum(stratum_df$outcome == 0),
        losses       = sum(stratum_df$outcome == -1),
        n            = nrow(stratum_df),
        favorability = fav_score,
        ci_lower     = test_result$ci_lower,
        ci_upper     = test_result$ci_upper,
        p_value      = test_result$p_value
      )
    })
  })
}

format_length_stratified_table <- function(results, axis_name) {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95        = sprintf("[%.1f%%, %.1f%%]",
                             100 * ci_lower, 100 * ci_upper),
      p_value      = sprintf("%.4f", p_value)
    ) %>%
    select(vendor, stratum, wins, ties, losses, n,
           favorability, ci_95, p_value) %>%
    arrange(vendor, stratum) %>%
    kable(caption = sprintf(
      paste0("Length-stratified one-vs-rest favorability test ",
             "(pooled rendering, %s axis). Stratum = vendor's answer ",
             "longer/shorter than comparator."),
      axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility

```{r length_stratified_clinical}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "clinical_utility") %>%
  format_length_stratified_table("clinical utility")
```

## Accuracy

```{r length_stratified_accuracy}
run_length_stratified_analysis(qa_ratings_with_lengths, "accuracy") %>%
  format_length_stratified_table("accuracy")
```

## Authoritativeness

```{r length_stratified_authoritativeness}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "authoritativeness") %>%
  format_length_stratified_table("authoritativeness")
```

## Completeness

```{r length_stratified_completeness}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "completeness") %>%
  format_length_stratified_table("completeness")
```

## Verifiability

```{r length_stratified_verifiability}
run_length_stratified_analysis(qa_ratings_with_lengths,
                               "verifiability") %>%
  format_length_stratified_table("verifiability")
```


# Exploratory analysis: Davidson rankings with ties

We fit a Davidson (1970) tie model via `BradleyTerry2::GenDavidson()` —
a non-linear term plugged into `gnm::gnm()` on the trinomial (win /
tie / loss) expansion of the contests. Like the Rao-Kupper model, this
extends Bradley-Terry to handle ties; ties are modelled by a single
parameter $\delta_2$ that multiplies the geometric mean of the two
vendors' worths:

$$
P(i \succ j) = \frac{\alpha_i}{\alpha_i + \alpha_j + \delta_2 \sqrt{\alpha_i \alpha_j}}, \qquad
P(\text{tie}) = \frac{\delta_2 \sqrt{\alpha_i \alpha_j}}{\alpha_i + \alpha_j + \delta_2 \sqrt{\alpha_i \alpha_j}}.
$$

Larger $\delta_2$ implies more ties; $\delta_2 = 0$ recovers
Bradley-Terry. BradleyTerry2 parameterises the tie via `tie.max`, the
logit of the maximum tie probability; the equivalent
$\delta_2 = 2\,\text{plogis}(\text{tie.max}) / (1 - \text{plogis}(\text{tie.max}))$.
Individual ability coefficients are not identifiable in the
non-linear term, so we use `gnm::getContrasts()` to obtain estimable
mean-centered log abilities $\log \alpha_i - \overline{\log \alpha}$
with valid SEs. We then report
$\tilde{\alpha}_i = \exp(\log \alpha_i - \overline{\log \alpha})$,
i.e. each vendor's worth as a ratio to the geometric-mean worth, with
Wald CIs $\exp(\widehat{\log \alpha}_i \pm 1.96\,\widehat{\mathrm{SE}}_i)$.
$\tilde{\alpha} > 1$ means above-average, $< 1$ means below-average.
All fits below are restricted to the `qa_text_only` and
`qa_text_citations` rendering modes (the `qa_full_render` condition is
excluded).

```{r davidson_helpers}
suppressMessages({
  library(BradleyTerry2)
  library(gnm)
})

fit_davidson <- function(df) {
  df      <- df %>% filter(!is.na(choice))
  vendors <- sort(unique(c(df$slot_a_vendor, df$slot_b_vendor)))
  win_a   <- !df$is_tie & df$winner == df$slot_a_vendor
  win_b   <- !df$is_tie & df$winner == df$slot_b_vendor

  expanded <- data.frame(
    match_id = factor(rep(seq_len(nrow(df)), each = 3)),
    player1  = factor(rep(df$slot_a_vendor, each = 3), levels = vendors),
    player2  = factor(rep(df$slot_b_vendor, each = 3), levels = vendors),
    result   = rep(c(1L, 0L, -1L), times = nrow(df)),
    count    = as.integer(c(t(cbind(win_a, df$is_tie, win_b))))
  )

  fit <- gnm(
    count ~ GenDavidson(result == 1, result == 0, result == -1,
                        player1, player2,
                        home.adv = NULL, tie.max = ~1) - 1,
    eliminate = match_id, family = poisson, data = expanded,
    verbose = FALSE
  )

  cc          <- coef(fit)
  tie_idx     <- grep("tie\\.|home\\.adv", names(cc))
  ability_idx <- setdiff(seq_along(cc), tie_idx)

  tie_max   <- as.numeric(cc[tie_idx[1]])
  p_tie_max <- plogis(tie_max)
  delta_2   <- 2 * p_tie_max / (1 - p_tie_max)

  # gnm does not report SEs for individual ability coefficients in
  # non-linear terms (only contrasts are identified). getContrasts gives
  # each log α centered at the mean log α, with proper SEs.
  qv <- getContrasts(fit, ability_idx)$qvframe

  log_alpha <- setNames(rep(NA_real_, length(vendors)), vendors)
  log_se    <- setNames(rep(NA_real_, length(vendors)), vendors)
  for (v in vendors) {
    idx <- which(endsWith(rownames(qv), v))
    if (length(idx) > 0) {
      log_alpha[v] <- qv$estimate[idx[1]]
      log_se[v]    <- qv$SE[idx[1]]
    }
  }

  z <- qnorm(0.975)

  list(
    vendors       = vendors,
    worth         = unname(exp(log_alpha)),
    worth_lower   = unname(exp(log_alpha - z * log_se)),
    worth_upper   = unname(exp(log_alpha + z * log_se)),
    delta_2       = delta_2,
    n_comparisons = nrow(df),
    n_questions   = length(unique(df$question_id)),
    fit           = fit
  )
}

format_davidson_table <- function(dav, caption) {
  caption_full <- sprintf(
    "%s — δ₂ = %.2f; N = %d comparisons across %d questions",
    caption, dav$delta_2, dav$n_comparisons, dav$n_questions
  )
  tibble(
    vendor   = dav$vendors,
    worth    = dav$worth,
    ci_lower = dav$worth_lower,
    ci_upper = dav$worth_upper
  ) %>%
    arrange(desc(worth)) %>%
    mutate(
      rank  = row_number(),
      worth = sprintf("%.3f [%.3f, %.3f]", worth, ci_lower, ci_upper)
    ) %>%
    select(rank, vendor, worth) %>%
    kable(caption   = caption_full,
          col.names = c("Rank", "Vendor",
                        "Worth ratio α̃ [95% CI]")) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```


```{r davidson_per_axis, results='asis'}
for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  dav_axis <- fit_davidson(qa_ratings_included %>% filter(axis == ax))
  print(format_davidson_table(
    dav_axis,
    sprintf("Davidson rankings on axis = %s", ax)
  ))
}
```

# Save everything
```{r save_outputs}
# Save workspace for R users
rdata_path <- file.path(params$output_dir, paste0(params$output_name, ".RData"))
save.image(file = rdata_path)
message("Saved RData to: ", rdata_path)
```