---
title: "Descriptives only"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)

source("data_helpers.R")
```

# Load data

Vendor identities are blinded by `blind_vendors.R`, which writes the
`blinded_*.csv` files read below and a `secret_mapping.txt` file (kept
outside the analysis pipeline). Run `Rscript blind_vendors.R` before
knitting this document.

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users)
```

# Table 1: Cohort summary

```{r table1}
completed_assignments <- assignments %>% filter(!is.na(completed_at))

n_questions_total  <- n_distinct(questions$id)
n_questions_used   <- n_distinct(qa_ratings$question_id)
n_graders_assigned <- n_distinct(assignments$user_id)
n_graders_rated    <- n_distinct(qa_ratings$user_id)
n_specialties_rated      <- n_distinct(qa_ratings$specialty)
n_specialties      <- n_distinct(questions$specialty)

per_user_completed <- completed_assignments %>%
  count(user_id, name = "n_completed")

avg_completed_per_user <- per_user_completed %>%
  summarise(
    mean_n = mean(n_completed),
    sd_n   = sd(n_completed),
    median_n = median(n_completed),
    min_n = min(n_completed),
    max_n = max(n_completed)
  )

submission_time_stats <- qa_ratings %>%
  filter(!is.na(submission_time_seconds)) %>%
  summarise(
    mean_t   = mean(submission_time_seconds),
    sd_t     = sd(submission_time_seconds),
    median_t = median(submission_time_seconds),
    min_t    = min(submission_time_seconds),
    max_t    = max(submission_time_seconds)
  )

users_per_subversion <- qa_ratings %>%
  distinct(user_id) %>%
  inner_join(users %>% select(id, subversion),
             by = c("user_id" = "id")) %>%
  count(subversion, name = "n_users") %>%
  arrange(subversion)

subversion_summary <- paste(
  sprintf("%s: %d", users_per_subversion$subversion,
          users_per_subversion$n_users),
  collapse = "; "
)

users_with_ratings <- unique(qa_ratings$user_id)

per_user_rate <- users %>%
  filter(id %in% users_with_ratings) %>%
  left_join(per_user_completed, by = c("id" = "user_id")) %>%
  mutate(
    n_completed = coalesce(n_completed, 0L),
    completion_rate = n_completed / required_task_count
  )

completion_rate_stats <- per_user_rate %>%
  summarise(
    mean_r = mean(completion_rate),
    sd_r   = sd(completion_rate)
  )

table1 <- tibble(
  Metric = c(
    "Total questions in question bank",
    "Total unique specialties in question bank",
    "Questions with ≥1 rating",
    "Specialties with ≥1 rating",
    "Graders who submitted ≥1 rating",
    "Graders with ≥1 rating per rendering mode",
    "Completed assignments per grader (mean ± SD)",
    "Completed assignments per grader (median [min, max])",
    "Survey completion rate per grader (mean ± SD)",
    "Submission time, seconds (mean ± SD)",
    "Submission time, seconds (median [min, max])"
  ),
  Value = c(
    n_questions_total,
    n_specialties,
    n_questions_used,
    n_specialties_rated,
    n_graders_rated,
    subversion_summary,
    sprintf("%.1f ± %.1f",
            avg_completed_per_user$mean_n,
            avg_completed_per_user$sd_n),
    sprintf("%d [%d, %d]",
            as.integer(avg_completed_per_user$median_n),
            as.integer(avg_completed_per_user$min_n),
            as.integer(avg_completed_per_user$max_n)),
    sprintf("%.1f%% ± %.1f%%",
            100 * completion_rate_stats$mean_r,
            100 * completion_rate_stats$sd_r),
    sprintf("%.1f ± %.1f",
            submission_time_stats$mean_t,
            submission_time_stats$sd_t),
    sprintf("%d [%d, %d]",
            as.integer(submission_time_stats$median_t),
            as.integer(submission_time_stats$min_t),
            as.integer(submission_time_stats$max_t))
  )
)

kable(table1, caption = "Table 1. Pilot cohort summary.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Submission time distribution

```{r submission_time_plot}
qa_ratings %>%
  filter(!is.na(submission_time_seconds)) %>%
  ggplot(aes(x = submission_time_seconds)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  scale_x_log10(labels = scales::comma) +
  labs(
    x = "Submission time (seconds, log scale)",
    y = "Number of ratings",
    title = "Distribution of submission time per rating"
  ) +
  theme_minimal()
```

## Questions per specialty

```{r specialty}
questions %>%
  count(specialty, name = "n_questions") %>%
  arrange(desc(n_questions)) %>%
  mutate(pct = sprintf("%.1f%%", 100 * n_questions / sum(n_questions))) %>%
  kable(caption = "Number of questions per specialty (full question bank).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

```{r specialty_used}
qa_ratings %>%
  distinct(question_id, specialty) %>%
  count(specialty, name = "n_questions_rated") %>%
  arrange(desc(n_questions_rated)) %>%
  kable(caption = "Number of rated questions per specialty (pilot).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```