---
title: "Descriptives only"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)

source("data_helpers.R")
```

# Load data

Vendor identities are blinded by `blind_vendors.R`, which writes the
`blinded_*.csv` files read below and a `secret_mapping.csv` file (kept
outside the analysis pipeline). Run `Rscript blind_vendors.R` before
knitting this document.

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
prior_users_path <- file.path(data_dir, "blinded_prior_users.csv")
prior_users <- if (file.exists(prior_users_path)) {
  read_csv(prior_users_path)
} else {
  NULL
}
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users, prior_users = prior_users)
```

# Table 1: Descriptive Statistics

```{r table1}
n_questions_total <- n_distinct(questions$id)
n_specialties     <- n_distinct(questions$specialty)

assignments_in_queue <- assignments %>%
  { if (!is.null(prior_users)) filter(., !user_id %in% prior_users$id) else . } %>%
  count(user_id, name = "n_assigned")

summarise_cohort <- function(qa_ratings_subset, users, assignments_in_queue) {
  per_user_completed <- qa_ratings_subset %>%
    distinct(user_id, qa_assignment_id) %>%
    count(user_id, name = "n_completed")

  avg_completed_per_user <- per_user_completed %>%
    summarise(
      mean_n   = mean(n_completed),
      sd_n     = sd(n_completed),
      median_n = median(n_completed),
      min_n    = min(n_completed),
      max_n    = max(n_completed)
    )

  submission_time_stats <- qa_ratings_subset %>%
    filter(!is.na(submission_time_seconds)) %>%
    summarise(
      mean_t   = mean(submission_time_seconds),
      sd_t     = sd(submission_time_seconds),
      median_t = median(submission_time_seconds),
      min_t    = min(submission_time_seconds),
      max_t    = max(submission_time_seconds)
    )

  ratings_per_question_stats <- qa_ratings_subset %>%
    count(question_id, name = "n_ratings") %>%
    summarise(
      mean_n   = mean(n_ratings),
      sd_n     = sd(n_ratings),
      median_n = median(n_ratings),
      min_n    = min(n_ratings),
      max_n    = max(n_ratings)
    )

  users_per_subversion <- qa_ratings_subset %>%
    distinct(user_id) %>%
    inner_join(users %>% select(id, subversion),
               by = c("user_id" = "id")) %>%
    count(subversion, name = "n_users") %>%
    arrange(subversion)

  subversion_summary <- paste(
    sprintf("%s: %d", users_per_subversion$subversion,
            users_per_subversion$n_users),
    collapse = "; "
  )

  per_user_rate <- users %>%
    filter(id %in% unique(qa_ratings_subset$user_id)) %>%
    left_join(per_user_completed, by = c("id" = "user_id")) %>%
    left_join(assignments_in_queue, by = c("id" = "user_id")) %>%
    mutate(
      n_completed     = coalesce(n_completed, 0L),
      completion_rate = n_completed / n_assigned
    )

  completion_rate_stats <- per_user_rate %>%
    summarise(
      mean_r = mean(completion_rate),
      sd_r   = sd(completion_rate)
    )

  c(
    n_questions_total,
    n_specialties,
    n_distinct(qa_ratings_subset$question_id),
    sprintf("%.1f ± %.1f",
            ratings_per_question_stats$mean_n,
            ratings_per_question_stats$sd_n),
    sprintf("%d [%d, %d]",
            as.integer(ratings_per_question_stats$median_n),
            as.integer(ratings_per_question_stats$min_n),
            as.integer(ratings_per_question_stats$max_n)),
    n_distinct(qa_ratings_subset$specialty),
    n_distinct(qa_ratings_subset$user_id),
    subversion_summary,
    sprintf("%.1f ± %.1f",
            avg_completed_per_user$mean_n,
            avg_completed_per_user$sd_n),
    sprintf("%d [%d, %d]",
            as.integer(avg_completed_per_user$median_n),
            as.integer(avg_completed_per_user$min_n),
            as.integer(avg_completed_per_user$max_n)),
    sprintf("%.1f%% ± %.1f%%",
            100 * completion_rate_stats$mean_r,
            100 * completion_rate_stats$sd_r),
    sprintf("%.1f ± %.1f",
            submission_time_stats$mean_t,
            submission_time_stats$sd_t),
    sprintf("%d [%d, %d]",
            as.integer(submission_time_stats$median_t),
            as.integer(submission_time_stats$min_t),
            as.integer(submission_time_stats$max_t))
  )
}

qa_ratings_passing <- apply_inclusion_criteria(qa_ratings)

table1 <- tibble(
  Metric = c(
    "Total questions in question bank",
    "Total unique specialties in question bank",
    "Questions with ≥1 rating",
    "Ratings per question (mean ± SD)",
    "Ratings per question (median [min, max])",
    "Specialties with ≥1 rating",
    "Graders who submitted ≥1 rating",
    "Graders with ≥1 rating per rendering mode",
    "Completed assignments per grader (mean ± SD)",
    "Completed assignments per grader (median [min, max])",
    "Survey completion rate per grader (mean ± SD)",
    "Submission time, seconds (mean ± SD)",
    "Submission time, seconds (median [min, max])"
  ),
  `All ratings` = summarise_cohort(qa_ratings, users, assignments_in_queue),
  `Submission time ≥ 10s` = summarise_cohort(qa_ratings_passing, users, assignments_in_queue)
)

kable(table1, caption = "Table 1. Pilot cohort summary, with and without submission-time exclusion.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Submission time distribution

```{r submission_time_plot}
qa_ratings %>%
  filter(!is.na(submission_time_seconds)) %>%
  ggplot(aes(x = submission_time_seconds)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  scale_x_log10(labels = scales::comma) +
  labs(
    x = "Submission time (seconds, log scale)",
    y = "Number of ratings",
    title = "Distribution of submission time per rating"
  ) +
  theme_minimal()
```

## Questions per specialty

```{r specialty}
specialty_counts <- questions %>%
  count(specialty, name = "n_questions") %>%
  arrange(desc(n_questions)) %>%
  mutate(pct = sprintf("%.1f%%", 100 * n_questions / sum(n_questions)))

specialty_counts %>%
  kable(caption = sprintf(
    "Number of questions per specialty (full question bank; %d unique specialties).",
    nrow(specialty_counts)
  )) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

```{r specialty_used}
specialty_summary <- qa_ratings %>%
  group_by(specialty) %>%
  summarise(
    n_questions_rated = n_distinct(question_id),
    n_graders = n_distinct(user_id),
    .groups = "drop"
  ) %>%
  arrange(desc(n_questions_rated))

specialty_total <- tibble(
  specialty = "Total",
  n_questions_rated = n_distinct(qa_ratings$question_id),
  n_graders = n_distinct(qa_ratings$user_id)
)

bind_rows(specialty_summary, specialty_total) %>%
  kable(caption = "Number of rated questions per specialty (pilot).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```