---
title: "Descriptives only"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)

source("data_helpers.R")
```

# Load data

Vendor identities are blinded by `blind_vendors.R`, which writes the
`blinded_*.csv` files read below and a `secret_mapping.csv` file (kept
outside the analysis pipeline). Run `Rscript blind_vendors.R` before
knitting this document.

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
qa_answers   <- read_csv(file.path(data_dir, "blinded_qa_answers.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
prior_users_path <- file.path(data_dir, "blinded_prior_users.csv")
prior_users <- if (file.exists(prior_users_path)) {
  read_csv(prior_users_path)
} else {
  NULL
}
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users, prior_users = prior_users)
```

# Table 1: Descriptive Statistics

```{r table1}
n_questions_total <- n_distinct(questions$id)
n_specialties     <- n_distinct(questions$specialty)

assignments_in_queue <- assignments %>%
  { if (!is.null(prior_users)) filter(., !user_id %in% prior_users$id) else . } %>%
  count(user_id, name = "n_assigned")

summarise_cohort <- function(qa_ratings_subset, users, assignments_in_queue) {
  mean_ratings_per_axis <- qa_ratings_subset %>%
    count(axis, name = "n_ratings") %>%
    summarise(mean_n = mean(n_ratings)) %>%
    pull(mean_n)

  ratings_per_vendor <- qa_ratings_subset %>%
    distinct(qa_assignment_id, slot_a_vendor, slot_b_vendor) %>%
    pivot_longer(c(slot_a_vendor, slot_b_vendor),
                 names_to = "slot", values_to = "vendor") %>%
    count(vendor, name = "n_comparisons") %>%
    arrange(vendor)

  ratings_per_vendor_summary <- paste(
    sprintf("%s: %d", ratings_per_vendor$vendor,
            ratings_per_vendor$n_comparisons),
    collapse = "; "
  )

  questions_per_vendor <- qa_ratings_subset %>%
    distinct(qa_assignment_id, question_id, slot_a_vendor, slot_b_vendor) %>%
    pivot_longer(c(slot_a_vendor, slot_b_vendor),
                 names_to = "slot", values_to = "vendor") %>%
    distinct(vendor, question_id) %>%
    count(vendor, name = "n_questions") %>%
    arrange(vendor)

  questions_per_vendor_summary <- paste(
    sprintf("%s: %d", questions_per_vendor$vendor,
            questions_per_vendor$n_questions),
    collapse = "; "
  )

  per_user_completed <- qa_ratings_subset %>%
    distinct(user_id, qa_assignment_id) %>%
    count(user_id, name = "n_completed")

  avg_completed_per_user <- per_user_completed %>%
    summarise(
      mean_n   = mean(n_completed),
      sd_n     = sd(n_completed),
      median_n = median(n_completed),
      min_n    = min(n_completed),
      max_n    = max(n_completed)
    )

  submission_time_stats <- qa_ratings_subset %>%
    filter(!is.na(submission_time_seconds)) %>%
    summarise(
      mean_t   = mean(submission_time_seconds),
      sd_t     = sd(submission_time_seconds),
      median_t = median(submission_time_seconds),
      min_t    = min(submission_time_seconds),
      max_t    = max(submission_time_seconds)
    )

  ratings_per_question_stats <- qa_ratings_subset %>%
    count(question_id, name = "n_ratings") %>%
    summarise(
      mean_n   = mean(n_ratings),
      sd_n     = sd(n_ratings),
      median_n = median(n_ratings),
      min_n    = min(n_ratings),
      max_n    = max(n_ratings)
    )

  users_per_subversion <- qa_ratings_subset %>%
    distinct(user_id) %>%
    inner_join(users %>% select(id, subversion),
               by = c("user_id" = "id")) %>%
    count(subversion, name = "n_users") %>%
    arrange(subversion)

  subversion_summary <- paste(
    sprintf("%s: %d", users_per_subversion$subversion,
            users_per_subversion$n_users),
    collapse = "; "
  )

  per_user_rate <- users %>%
    filter(id %in% unique(qa_ratings_subset$user_id)) %>%
    left_join(per_user_completed, by = c("id" = "user_id")) %>%
    left_join(assignments_in_queue, by = c("id" = "user_id")) %>%
    mutate(
      n_completed     = coalesce(n_completed, 0L),
      completion_rate = n_completed / n_assigned
    )

  completion_rate_stats <- per_user_rate %>%
    summarise(
      mean_r = mean(completion_rate),
      sd_r   = sd(completion_rate)
    )

  c(
    n_questions_total,
    n_specialties,
    n_distinct(qa_ratings_subset$question_id),
    sprintf("%.0f", mean_ratings_per_axis),
    ratings_per_vendor_summary,
    questions_per_vendor_summary,
    sprintf("%.1f ± %.1f",
            ratings_per_question_stats$mean_n,
            ratings_per_question_stats$sd_n),
    sprintf("%d [%d, %d]",
            as.integer(ratings_per_question_stats$median_n),
            as.integer(ratings_per_question_stats$min_n),
            as.integer(ratings_per_question_stats$max_n)),
    n_distinct(qa_ratings_subset$specialty),
    n_distinct(qa_ratings_subset$user_id),
    subversion_summary,
    sprintf("%.1f ± %.1f",
            avg_completed_per_user$mean_n,
            avg_completed_per_user$sd_n),
    sprintf("%d [%d, %d]",
            as.integer(avg_completed_per_user$median_n),
            as.integer(avg_completed_per_user$min_n),
            as.integer(avg_completed_per_user$max_n)),
    sprintf("%.1f%% ± %.1f%%",
            100 * completion_rate_stats$mean_r,
            100 * completion_rate_stats$sd_r),
    sprintf("%.1f ± %.1f",
            submission_time_stats$mean_t,
            submission_time_stats$sd_t),
    sprintf("%d [%d, %d]",
            as.integer(submission_time_stats$median_t),
            as.integer(submission_time_stats$min_t),
            as.integer(submission_time_stats$max_t))
  )
}

qa_ratings_passing <- apply_inclusion_criteria(qa_ratings)

table1 <- tibble(
  Metric = c(
    "Total questions in question bank",
    "Total unique specialties in question bank",
    "Questions with ≥1 rating",
    "Total ratings collected (mean across evaluation dimensions)",
    "Ratings collected per blinded vendor (comparisons)",
    "Unique questions evaluated per blinded vendor",
    "Ratings per question (mean ± SD)",
    "Ratings per question (median [min, max])",
    "Specialties with ≥1 rating",
    "Graders who submitted ≥1 rating",
    "Graders with ≥1 rating per rendering mode",
    "Completed assignments per grader (mean ± SD)",
    "Completed assignments per grader (median [min, max])",
    "Survey completion rate per grader (mean ± SD)",
    "Submission time, seconds (mean ± SD)",
    "Submission time, seconds (median [min, max])"
  ),
  `All ratings` = summarise_cohort(qa_ratings, users, assignments_in_queue),
  `Submission time ≥ 10s` = summarise_cohort(qa_ratings_passing, users, assignments_in_queue)
)

kable(table1, caption = "Table 1. Pilot cohort summary, with and without submission-time exclusion.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Submission time distribution

```{r submission_time_plot}
qa_ratings %>%
  filter(!is.na(submission_time_seconds)) %>%
  ggplot(aes(x = submission_time_seconds)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  scale_x_log10(labels = scales::comma) +
  labs(
    x = "Submission time (seconds, log scale)",
    y = "Number of ratings",
    title = "Distribution of submission time per rating"
  ) +
  theme_minimal()
```

## Questions per specialty

```{r specialty}
specialty_counts <- questions %>%
  count(specialty, name = "n_questions") %>%
  arrange(desc(n_questions)) %>%
  mutate(pct = sprintf("%.1f%%", 100 * n_questions / sum(n_questions)))

specialty_counts %>%
  kable(caption = sprintf(
    "Number of questions per specialty (full question bank; %d unique specialties).",
    nrow(specialty_counts)
  )) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

<!-- ```{r specialty_list}
paste0(unique(questions$specialty), collapse=", ")
``` -->

```{r specialty_used}
specialty_summary <- qa_ratings %>%
  group_by(specialty) %>%
  summarise(
    n_questions_rated = n_distinct(question_id),
    n_graders = n_distinct(user_id),
    .groups = "drop"
  ) %>%
  arrange(desc(n_questions_rated))

specialty_total <- tibble(
  specialty = "Total",
  n_questions_rated = n_distinct(qa_ratings$question_id),
  n_graders = n_distinct(qa_ratings$user_id)
)

bind_rows(specialty_summary, specialty_total) %>%
  kable(caption = "Number of rated questions per specialty (pilot).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Answer length distribution by vendor

Each rated assignment shows the grader two answers (slots A and B). The
plots below describe the character lengths of those rated answers,
stratified by the blinded vendor that produced them.

```{r answer_lengths_rated}
rated_answer_lengths <- qa_ratings %>%
  select(question_id, slot_a_vendor, slot_b_vendor) %>%
  pivot_longer(c(slot_a_vendor, slot_b_vendor),
               names_to = "slot", values_to = "vendor") %>%
  inner_join(qa_answers %>% select(question_id, vendor, answer_len),
             by = c("question_id", "vendor")) %>%
  filter(!is.na(answer_len))
```

```{r answer_lengths_table}
rated_answer_lengths %>%
  group_by(vendor) %>%
  summarise(
    n_rated_answers = n(),
    mean_len   = mean(answer_len),
    sd_len     = sd(answer_len),
    median_len = median(answer_len),
    min_len    = min(answer_len),
    max_len    = max(answer_len),
    .groups = "drop"
  ) %>%
  arrange(vendor) %>%
  mutate(
    `Mean ± SD` = sprintf("%.0f ± %.0f", mean_len, sd_len),
    `Median [min, max]` = sprintf("%.0f [%.0f, %.0f]",
                                  median_len, min_len, max_len)
  ) %>%
  select(Vendor = vendor, `Rated answers (n)` = n_rated_answers,
         `Mean ± SD`, `Median [min, max]`) %>%
  kable(caption = "Answer length (characters) per blinded vendor, across rated assignments.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

```{r answer_lengths_histogram}
ggplot(rated_answer_lengths, aes(x = answer_len, fill = vendor)) +
  geom_histogram(bins = 40, color = "white", show.legend = FALSE) +
  facet_wrap(~ vendor, scales = "free_y") +
  scale_x_log10(labels = scales::comma) +
  labs(
    x = "Answer length (characters, log scale)",
    y = "Number of rated answers",
    title = "Distribution of rated answer lengths, by blinded vendor"
  ) +
  theme_minimal()
```

```{r answer_lengths_boxplot}
ggplot(rated_answer_lengths, aes(x = vendor, y = answer_len, fill = vendor)) +
  geom_boxplot(show.legend = FALSE, outlier.alpha = 0.4) +
  scale_y_log10(labels = scales::comma) +
  labs(
    x = "Blinded vendor",
    y = "Answer length (characters, log scale)",
    title = "Rated answer length by blinded vendor"
  ) +
  theme_minimal()
```

## Length-based win-rate matrix

For each ordered pair of vendors (row, column), the cell reports the
fraction of questions on which the row vendor's answer is strictly
longer (in characters) than the column vendor's. The denominator is the
number of questions for which both vendors have an answer length on
record. Diagonal cells are blank.

```{r length_win_matrix}
answer_len_lookup <- qa_answers %>%
  filter(!is.na(answer_len)) %>%
  select(question_id, vendor, answer_len)

vendor_levels <- sort(unique(answer_len_lookup$vendor))

len_a_lookup <- answer_len_lookup %>%
  rename(slot_a_vendor = vendor, len_a = answer_len)
len_b_lookup <- answer_len_lookup %>%
  rename(slot_b_vendor = vendor, len_b = answer_len)

rated_comparisons <- qa_ratings %>%
  distinct(qa_assignment_id, question_id, slot_a_vendor, slot_b_vendor) %>%
  inner_join(len_a_lookup, by = c("question_id", "slot_a_vendor")) %>%
  inner_join(len_b_lookup, by = c("question_id", "slot_b_vendor"))

length_pairs <- bind_rows(
  rated_comparisons %>%
    transmute(vendor_row = slot_a_vendor, vendor_col = slot_b_vendor,
              row_longer = len_a > len_b),
  rated_comparisons %>%
    transmute(vendor_row = slot_b_vendor, vendor_col = slot_a_vendor,
              row_longer = len_b > len_a)
) %>%
  group_by(vendor_row, vendor_col) %>%
  summarise(
    n_comparisons = n(),
    n_row_longer = sum(row_longer),
    win_rate = n_row_longer / n_comparisons,
    .groups = "drop"
  ) %>%
  mutate(
    vendor_row = factor(vendor_row, levels = vendor_levels),
    vendor_col = factor(vendor_col, levels = vendor_levels)
  )
```

```{r length_win_matrix_table}
length_pairs %>%
  mutate(cell = sprintf("%.1f%% (%d/%d)",
                        100 * win_rate, n_row_longer, n_comparisons)) %>%
  select(vendor_row, vendor_col, cell) %>%
  pivot_wider(names_from = vendor_col, values_from = cell, values_fill = "—",
              names_sort = TRUE) %>%
  arrange(vendor_row) %>%
  rename(`Row vendor` = vendor_row) %>%
  kable(caption = paste(
    "Length-based win rate among rated comparisons:",
    "P(row vendor's answer is longer than column vendor's),",
    "with (wins / rated comparisons)."
  )) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```