---
title: "Pilot Data: Descriptive Statistics & Win Rates"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(lme4)
library(broom.mixed)
```

# Load data

```{r load}
data_dir <- "pilot_data"

assignments <- read_csv(file.path(data_dir, "qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "ratings.csv"))
users       <- read_csv(file.path(data_dir, "users.csv"))
```

# Mask vendor identities

All vendor names are replaced with masked labels (A, B, C, D) for the
remainder of the analysis. The mapping is intentionally scrambled and lives
only in this chunk.

```{r mask, echo = FALSE}
vendor_mapping <- c(
  "gpt-5.5"        = "A",
  "openevidence"   = "B",
  "claude-opus-4-8"= "C",
  "gemini-3.1-pro" = "D"
)

mask_vendor <- function(x) unname(vendor_mapping[x])

assignments <- assignments %>%
  mutate(
    slot_a_vendor = mask_vendor(slot_a_provider),
    slot_b_vendor = mask_vendor(slot_b_provider)
  ) %>%
  select(-slot_a_provider, -slot_b_provider)
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- ratings %>%
  filter(!is.na(qa_assignment_id)) %>%
  inner_join(
    assignments %>%
      select(assignment_id = id, user_id, question_id,
             slot_a_vendor, slot_b_vendor, completed_at),
    by = c("qa_assignment_id" = "assignment_id")
  ) %>%
  inner_join(
    questions %>% select(question_id = id, specialty),
    by = "question_id"
  ) %>%
  inner_join(
    users %>% select(user_id = id, subversion),
    by = "user_id"
  ) %>%
  mutate(
    winner = case_when(
      choice %in% c("strongly_a", "slightly_a") ~ slot_a_vendor,
      choice %in% c("strongly_b", "slightly_b") ~ slot_b_vendor,
      choice == "tie" ~ NA_character_
    ),
    loser = case_when(
      choice %in% c("strongly_a", "slightly_a") ~ slot_b_vendor,
      choice %in% c("strongly_b", "slightly_b") ~ slot_a_vendor,
      choice == "tie" ~ NA_character_
    ),
    is_tie = choice == "tie"
  )
```

# Table 1: Cohort summary

```{r table1}
completed_assignments <- assignments %>% filter(!is.na(completed_at))

n_questions_total  <- n_distinct(questions$id)
n_questions_used   <- n_distinct(qa_ratings$question_id)
n_graders_assigned <- n_distinct(assignments$user_id)
n_graders_rated    <- n_distinct(qa_ratings$user_id)
n_specialties      <- n_distinct(questions$specialty)

per_user_completed <- completed_assignments %>%
  count(user_id, name = "n_completed")

avg_completed_per_user <- per_user_completed %>%
  summarise(
    mean_n = mean(n_completed),
    sd_n   = sd(n_completed),
    median_n = median(n_completed),
    min_n = min(n_completed),
    max_n = max(n_completed)
  )

submission_time_stats <- qa_ratings %>%
  filter(!is.na(submission_time_seconds)) %>%
  summarise(
    mean_t   = mean(submission_time_seconds),
    sd_t     = sd(submission_time_seconds),
    median_t = median(submission_time_seconds),
    min_t    = min(submission_time_seconds),
    max_t    = max(submission_time_seconds)
  )

users_per_subversion <- assignments %>%
  inner_join(users %>% select(id, subversion),
             by = c("user_id" = "id")) %>%
  distinct(user_id, subversion) %>%
  count(subversion, name = "n_users") %>%
  arrange(subversion)

subversion_summary <- paste(
  sprintf("%s: %d", users_per_subversion$subversion,
          users_per_subversion$n_users),
  collapse = "; "
)

users_with_ratings <- unique(qa_ratings$user_id)

per_user_rate <- users %>%
  filter(id %in% users_with_ratings) %>%
  left_join(per_user_completed, by = c("id" = "user_id")) %>%
  mutate(
    n_completed = coalesce(n_completed, 0L),
    completion_rate = n_completed / required_task_count
  )

completion_rate_stats <- per_user_rate %>%
  summarise(
    mean_r = mean(completion_rate),
    sd_r   = sd(completion_rate)
  )

table1 <- tibble(
  Metric = c(
    "Total questions in question bank",
    "Questions used in pilot (≥1 rating)",
    "Specialties represented",
    "Graders assigned ≥1 question",
    "Graders who submitted ≥1 rating",
    "Graders assigned per rendering mode",
    "Completed assignments per grader (mean ± SD)",
    "Completed assignments per grader (median [min, max])",
    "Survey completion rate per grader (mean ± SD)",
    "Submission time, seconds (mean ± SD)",
    "Submission time, seconds (median [min, max])"
  ),
  Value = c(
    n_questions_total,
    n_questions_used,
    n_specialties,
    n_graders_assigned,
    n_graders_rated,
    subversion_summary,
    sprintf("%.1f ± %.1f",
            avg_completed_per_user$mean_n,
            avg_completed_per_user$sd_n),
    sprintf("%d [%d, %d]",
            as.integer(avg_completed_per_user$median_n),
            as.integer(avg_completed_per_user$min_n),
            as.integer(avg_completed_per_user$max_n)),
    sprintf("%.1f%% ± %.1f%%",
            100 * completion_rate_stats$mean_r,
            100 * completion_rate_stats$sd_r),
    sprintf("%.1f ± %.1f",
            submission_time_stats$mean_t,
            submission_time_stats$sd_t),
    sprintf("%d [%d, %d]",
            as.integer(submission_time_stats$median_t),
            as.integer(submission_time_stats$min_t),
            as.integer(submission_time_stats$max_t))
  )
)

kable(table1, caption = "Table 1. Pilot cohort summary.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Submission time distribution

```{r submission_time_plot}
qa_ratings %>%
  filter(!is.na(submission_time_seconds)) %>%
  ggplot(aes(x = submission_time_seconds)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white") +
  scale_x_continuous(labels = scales::comma) +
  labs(
    x = "Submission time (seconds)",
    y = "Number of ratings",
    title = "Distribution of submission time per rating"
  ) +
  theme_minimal()
```

## Questions per specialty

```{r specialty}
questions %>%
  count(specialty, name = "n_questions") %>%
  arrange(desc(n_questions)) %>%
  mutate(pct = sprintf("%.1f%%", 100 * n_questions / sum(n_questions))) %>%
  kable(caption = "Number of questions per specialty (full question bank).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

```{r specialty_used}
qa_ratings %>%
  distinct(question_id, specialty) %>%
  count(specialty, name = "n_questions_rated") %>%
  arrange(desc(n_questions_rated)) %>%
  kable(caption = "Number of rated questions per specialty (pilot).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Win-rate tables

For every (assignment, axis) row we know which two vendors were compared
and the chosen winner. The matrix below reports, for each ordered pair
(row vendor vs. column vendor), the row vendor's win rate across all
axes. Ties are *included* in the denominator (and count as 0 wins for
both vendors), so cell[row, col] + cell[col, row] + tie-rate = 100%.

## Win-rate matrix

```{r winrate_matrix}
directed <- qa_ratings %>%
  filter(!is_tie) %>%
  transmute(row_vendor = winner, col_vendor = loser) %>%
  bind_rows(
    qa_ratings %>%
      filter(!is_tie) %>%
      transmute(row_vendor = loser, col_vendor = winner, .lost = TRUE)
  )

vendor_levels <- sort(unique(c(qa_ratings$slot_a_vendor, qa_ratings$slot_b_vendor)))

win_matrix <- qa_ratings %>%
  mutate(
    row_vendor = slot_a_vendor,
    col_vendor = slot_b_vendor,
    row_win = !is_tie & winner == slot_a_vendor
  ) %>%
  bind_rows(
    qa_ratings %>%
      mutate(
        row_vendor = slot_b_vendor,
        col_vendor = slot_a_vendor,
        row_win = !is_tie & winner == slot_b_vendor
      )
  ) %>%
  group_by(row_vendor, col_vendor) %>%
  summarise(
    n = n(),
    wins = sum(row_win),
    win_rate = wins / n,
    .groups = "drop"
  ) %>%
  mutate(cell = sprintf("%.1f%% (%d/%d)", 100 * win_rate, wins, n)) %>%
  select(row_vendor, col_vendor, cell) %>%
  pivot_wider(names_from = col_vendor, values_from = cell,
              values_fill = "—", names_sort = TRUE) %>%
  arrange(row_vendor)

kable(
  win_matrix,
  caption = "Row vendor's win rate vs. column vendor (ties included in denominator). Cell = win% (wins / all comparisons)."
) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Win-rate matrices per axis

Same construction as the pooled matrix, repeated within each rating axis.

```{r winrate_matrix_per_axis, results='asis'}
build_win_matrix <- function(df) {
  df %>%
    mutate(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_win = !is_tie & winner == slot_a_vendor
    ) %>%
    bind_rows(
      df %>%
        mutate(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_win = !is_tie & winner == slot_b_vendor
        )
    ) %>%
    group_by(row_vendor, col_vendor) %>%
    summarise(
      n = n(),
      wins = sum(row_win),
      win_rate = wins / n,
      .groups = "drop"
    ) %>%
    mutate(cell = sprintf("%.1f%% (%d/%d)", 100 * win_rate, wins, n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
              values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}

axes <- sort(unique(qa_ratings$axis))

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor on axis = %s ",
               "(ties included in denominator)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

## Win-rate matrices per rendering mode

Same construction as the pooled matrix, repeated within each rendering
mode (`subversion` in `users.csv`).

```{r winrate_matrix_per_subversion, results='asis'}
subversions <- sort(unique(qa_ratings$subversion))

for (sv in subversions) {
  cat(sprintf("\n\n### Rendering mode: %s\n\n", sv))
  mat <- build_win_matrix(qa_ratings %>% filter(subversion == sv))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Row vendor's win rate vs. column vendor for ",
               "rendering mode = %s (ties included in denominator)."),
        sv
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

# Win-margin tables

Same row-vs-column structure as the win-rate matrices, but each cell is
the mean number of Likert units the row vendor beat the column vendor
by. The Likert scale is `strongly_*` = 2 units, `slightly_*` = 1 unit,
`tie` = 0 units. Ties are *included* (they contribute 0 units). Cells
are mirrored: `cell[row, col] = -cell[col, row]`. Cell format:
`mean (net/N)` where `net` is the signed sum of units across all
comparisons in that cell.

```{r margin_helpers}
likert_units_a <- function(choice) {
  case_when(
    choice == "strongly_a" ~  2,
    choice == "slightly_a" ~  1,
    choice == "tie"        ~  0,
    choice == "slightly_b" ~ -1,
    choice == "strongly_b" ~ -2
  )
}

build_margin_matrix <- function(df) {
  df %>%
    mutate(units_a = likert_units_a(choice)) %>%
    transmute(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_units  = units_a
    ) %>%
    bind_rows(
      df %>%
        mutate(units_a = likert_units_a(choice)) %>%
        transmute(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_units  = -units_a
        )
    ) %>%
    group_by(row_vendor, col_vendor) %>%
    summarise(
      n = n(),
      net_units  = sum(row_units),
      mean_units = net_units / n,
      .groups = "drop"
    ) %>%
    mutate(cell = sprintf("%+.2f (%+d/%d)",
                          mean_units,
                          as.integer(net_units),
                          n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell,
                values_fill = "—", names_sort = TRUE) %>%
    arrange(row_vendor)
}
```

## Win-margin matrix

```{r margin_matrix}
kable(
  build_margin_matrix(qa_ratings),
  caption = paste0(
    "Mean Likert-unit margin the row vendor won by vs. the column ",
    "vendor (ties included as 0). Cell = mean (net/N)."
  )
) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "hover"))
```

## Win-margin matrices per axis

```{r margin_matrix_per_axis, results='asis'}
for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_margin_matrix(qa_ratings %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Mean Likert-unit margin row vendor won by vs. column ",
               "vendor on axis = %s (ties included)."),
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```

## Win-margin matrices per rendering mode

```{r margin_matrix_per_subversion, results='asis'}
for (sv in subversions) {
  cat(sprintf("\n\n### Rendering mode: %s\n\n", sv))
  mat <- build_margin_matrix(qa_ratings %>% filter(subversion == sv))
  print(
    kable(
      mat,
      caption = sprintf(
        paste0("Mean Likert-unit margin row vendor won by vs. column ",
               "vendor for rendering mode = %s (ties included)."),
        sv
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```