---
title: "Pilot Data: Descriptive Statistics & Win Rates"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Load data

```{r load}
data_dir <- "pilot_data"

assignments <- read_csv(file.path(data_dir, "qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "ratings.csv"))
users       <- read_csv(file.path(data_dir, "users.csv"))
```

# Mask vendor identities

All vendor names are replaced with masked labels (A, B, C, D) for the
remainder of the analysis. The mapping is intentionally scrambled and lives
only in this chunk.

```{r mask}
vendor_mapping <- c(
  "gpt-5.5"        = "A",
  "openevidence"   = "B",
  "claude-opus-4-8"= "C",
  "gemini-3.1-pro" = "D"
)

mask_vendor <- function(x) unname(vendor_mapping[x])

assignments <- assignments %>%
  mutate(
    slot_a_vendor = mask_vendor(slot_a_provider),
    slot_b_vendor = mask_vendor(slot_b_provider)
  ) %>%
  select(-slot_a_provider, -slot_b_provider)
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- ratings %>%
  filter(!is.na(qa_assignment_id)) %>%
  inner_join(
    assignments %>%
      select(assignment_id = id, user_id, question_id,
             slot_a_vendor, slot_b_vendor, completed_at),
    by = c("qa_assignment_id" = "assignment_id")
  ) %>%
  inner_join(
    questions %>% select(question_id = id, specialty),
    by = "question_id"
  ) %>%
  mutate(
    winner = case_when(
      choice %in% c("strongly_a", "slightly_a") ~ slot_a_vendor,
      choice %in% c("strongly_b", "slightly_b") ~ slot_b_vendor,
      choice == "tie" ~ NA_character_
    ),
    loser = case_when(
      choice %in% c("strongly_a", "slightly_a") ~ slot_b_vendor,
      choice %in% c("strongly_b", "slightly_b") ~ slot_a_vendor,
      choice == "tie" ~ NA_character_
    ),
    is_tie = choice == "tie"
  )
```

# Table 1: Cohort summary

```{r table1}
completed_assignments <- assignments %>% filter(!is.na(completed_at))

n_questions_total  <- n_distinct(questions$id)
n_questions_used   <- n_distinct(qa_ratings$question_id)
n_graders_assigned <- n_distinct(assignments$user_id)
n_graders_rated    <- n_distinct(qa_ratings$user_id)
n_specialties      <- n_distinct(questions$specialty)

per_user_completed <- completed_assignments %>%
  count(user_id, name = "n_completed")

avg_completed_per_user <- per_user_completed %>%
  summarise(
    mean_n = mean(n_completed),
    sd_n   = sd(n_completed),
    median_n = median(n_completed),
    min_n = min(n_completed),
    max_n = max(n_completed)
  )

table1 <- tibble(
  Metric = c(
    "Total questions in question bank",
    "Questions used in pilot (≥1 rating)",
    "Specialties represented",
    "Graders assigned ≥1 question",
    "Graders who submitted ≥1 rating",
    "Completed assignments per grader (mean ± SD)",
    "Completed assignments per grader (median [min, max])"
  ),
  Value = c(
    n_questions_total,
    n_questions_used,
    n_specialties,
    n_graders_assigned,
    n_graders_rated,
    sprintf("%.1f ± %.1f",
            avg_completed_per_user$mean_n,
            avg_completed_per_user$sd_n),
    sprintf("%d [%d, %d]",
            as.integer(avg_completed_per_user$median_n),
            as.integer(avg_completed_per_user$min_n),
            as.integer(avg_completed_per_user$max_n))
  )
)

kable(table1, caption = "Table 1. Pilot cohort summary.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Questions per specialty

```{r specialty}
questions %>%
  count(specialty, name = "n_questions") %>%
  arrange(desc(n_questions)) %>%
  mutate(pct = sprintf("%.1f%%", 100 * n_questions / sum(n_questions))) %>%
  kable(caption = "Number of questions per specialty (full question bank).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

```{r specialty_used}
qa_ratings %>%
  distinct(question_id, specialty) %>%
  count(specialty, name = "n_questions_rated") %>%
  arrange(desc(n_questions_rated)) %>%
  kable(caption = "Number of rated questions per specialty (pilot).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Win-rate tables

For every (assignment, axis) row we know which two vendors were compared
and the chosen winner. The matrix below reports, for each ordered pair
(row vendor vs. column vendor), the row vendor's win rate across all
axes. Ties are reported separately.

## Win-rate matrix

```{r winrate_matrix}
directed <- qa_ratings %>%
  filter(!is_tie) %>%
  transmute(row_vendor = winner, col_vendor = loser) %>%
  bind_rows(
    qa_ratings %>%
      filter(!is_tie) %>%
      transmute(row_vendor = loser, col_vendor = winner, .lost = TRUE)
  )

vendor_levels <- sort(unique(c(qa_ratings$slot_a_vendor, qa_ratings$slot_b_vendor)))

win_matrix <- qa_ratings %>%
  filter(!is_tie) %>%
  mutate(
    row_vendor = slot_a_vendor,
    col_vendor = slot_b_vendor,
    row_win = winner == slot_a_vendor
  ) %>%
  bind_rows(
    qa_ratings %>%
      filter(!is_tie) %>%
      mutate(
        row_vendor = slot_b_vendor,
        col_vendor = slot_a_vendor,
        row_win = winner == slot_b_vendor
      )
  ) %>%
  group_by(row_vendor, col_vendor) %>%
  summarise(
    n = n(),
    wins = sum(row_win),
    win_rate = wins / n,
    .groups = "drop"
  ) %>%
  mutate(cell = sprintf("%.1f%% (%d/%d)", 100 * win_rate, wins, n)) %>%
  select(row_vendor, col_vendor, cell) %>%
  pivot_wider(names_from = col_vendor, values_from = cell, values_fill = "—") %>%
  arrange(row_vendor)

kable(
  win_matrix,
  caption = "Row vendor's win rate vs. column vendor (ties excluded). Cell = win% (wins / non-tie comparisons)."
) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Win-rate matrices per axis

Same construction as the pooled matrix, repeated within each rating axis.

```{r winrate_matrix_per_axis, results='asis'}
build_win_matrix <- function(df) {
  df %>%
    filter(!is_tie) %>%
    mutate(
      row_vendor = slot_a_vendor,
      col_vendor = slot_b_vendor,
      row_win = winner == slot_a_vendor
    ) %>%
    bind_rows(
      df %>%
        filter(!is_tie) %>%
        mutate(
          row_vendor = slot_b_vendor,
          col_vendor = slot_a_vendor,
          row_win = winner == slot_b_vendor
        )
    ) %>%
    group_by(row_vendor, col_vendor) %>%
    summarise(
      n = n(),
      wins = sum(row_win),
      win_rate = wins / n,
      .groups = "drop"
    ) %>%
    mutate(cell = sprintf("%.1f%% (%d/%d)", 100 * win_rate, wins, n)) %>%
    select(row_vendor, col_vendor, cell) %>%
    pivot_wider(names_from = col_vendor, values_from = cell, values_fill = "—") %>%
    arrange(row_vendor)
}

axes <- sort(unique(qa_ratings$axis))

for (ax in axes) {
  cat(sprintf("\n\n### Axis: %s\n\n", ax))
  mat <- build_win_matrix(qa_ratings %>% filter(axis == ax))
  print(
    kable(
      mat,
      caption = sprintf(
        "Row vendor's win rate vs. column vendor on axis = %s (ties excluded).",
        ax
      )
    ) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "hover"))
  )
}
```
