---
title: "Statistical Analysis: Win Rates by Question Theme"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
  output_dir: null
  output_name: "exploratory_analysis"
  bootstrap_by: "question"
  drop_vendor: null
  user_stratification: null
  tagged_questions_path: "exp_themes/_output/prompt_v1/default/tagging_v1/tagged_questions.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)

source("favorability_test.R")
source("data_helpers.R")
```

# Analysis parameters

```{r params, echo=FALSE, results='asis'}
cat(sprintf("- **Data directory:** `%s`\n", params$data_dir))
cat(sprintf("- **Output directory:** `%s`\n", ifelse(is.null(params$output_dir), params$data_dir, params$output_dir)))
cat(sprintf("- **Bootstrap by:** %s\n", params$bootstrap_by))
cat(sprintf("- **User stratification:** %s\n", ifelse(is.null(params$user_stratification), "none", params$user_stratification)))
cat(sprintf("- **Drop vendor:** %s\n", ifelse(is.null(params$drop_vendor), "none", params$drop_vendor)))
cat(sprintf("- **Tagged questions path:** `%s`\n", params$tagged_questions_path))
```

# Load data

```{r load}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
user_registrations  <- read_csv(file.path(data_dir, "recruitment_combined.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
users <- get_eligible_users(users, user_registrations, user_stratification=params$user_stratification)
prior_users_path <- file.path(data_dir, "blinded_prior_users.csv")
prior_users <- if (file.exists(prior_users_path)) {
  read_csv(prior_users_path)
} else {
  NULL
}

tagged_questions <- read_csv(params$tagged_questions_path)
```

# Build the analytic dataset

Join ratings to assignments, attach question specialty and theme, and decode the
choice into a winner (vendor in slot A, vendor in slot B, or tie).

```{r analytic}
qa_ratings <- build_qa_ratings(ratings, assignments, questions, users, prior_users = prior_users)
qa_ratings <- apply_inclusion_criteria(qa_ratings)

question_themes <- tagged_questions %>%
  select(question_id = id, primary_question_type) %>%
  filter(!is.na(primary_question_type), primary_question_type != "")

qa_ratings <- qa_ratings %>%
  left_join(question_themes, by = "question_id")

n_missing_theme <- sum(is.na(qa_ratings$primary_question_type))
n_total <- nrow(qa_ratings)
cat(sprintf("Ratings with missing theme: %d / %d (%.1f%%)\n",
            n_missing_theme, n_total, 100 * n_missing_theme / n_total))

qa_ratings <- qa_ratings %>%
  filter(!is.na(primary_question_type))

if (!is.null(params$drop_vendor)) {
  stopifnot(params$drop_vendor %in% c("A", "B", "C", "D"))
  qa_ratings <- qa_ratings %>%
    filter(slot_a_vendor != params$drop_vendor,
           slot_b_vendor != params$drop_vendor)
  cat(sprintf("Dropped vendor %s from analysis\n", params$drop_vendor))
}
```

# Question theme distribution

```{r theme_distribution}
theme_counts <- qa_ratings %>%
  distinct(question_id, primary_question_type) %>%
  count(primary_question_type) %>%
  arrange(desc(n))

theme_counts %>%
  kable(caption = "Number of questions by primary theme",
        col.names = c("Theme", "N Questions")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Primary Analysis: Win difference tests by theme

One-vs-rest win difference (favorability score) test for each vendor stratified
by question theme, on the two primary outcome axes (accuracy and clinical_utility),
restricted to the text-only rendering condition.

```{r favorability_test_helpers}
prepare_one_vs_rest_df <- function(df, vendor) {
  df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      outcome = case_when(
        winner == vendor ~ 1L,
        loser == vendor ~ -1L,
        is_tie ~ 0L
      )
    ) %>%
    select(question_id, user_id, outcome)
}

run_favorability_analysis_by_theme <- function(df, axis_name, theme,
                                               subversion_filter = "qa_text_only",
                                               n_boot = 10000, seed = 42,
                                               bootstrap_by = params$bootstrap_by) {
  set.seed(seed)

  axis_df <- df %>%
    filter(subversion == subversion_filter,
           axis == axis_name,
           primary_question_type == theme)

  if (nrow(axis_df) == 0) return(NULL)

  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    test_df <- prepare_one_vs_rest_df(axis_df, v)

    if (nrow(test_df) == 0) return(NULL)

    fav_score <- get_favorability_score(test_df$outcome)
    test_result <- run_favorability_test(test_df, n_boot = n_boot, ci_level=0.95,
                                         bootstrap_by = bootstrap_by)

    tibble(
      theme = theme,
      vendor = v,
      wins = sum(test_df$outcome == 1),
      ties = sum(test_df$outcome == 0),
      losses = sum(test_df$outcome == -1),
      n = nrow(test_df),
      favorability = fav_score,
      ci_lower = test_result$ci_lower,
      ci_upper = test_result$ci_upper,
      p_value = test_result$p_value
    )
  })
}

format_favorability_table_by_theme <- function(results, axis_name,
                                               rendering = "text-only") {
  results %>%
    mutate(
      favorability = sprintf("%.1f%%", 100 * favorability),
      ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
      p_value = sprintf("%.4f", p_value)
    ) %>%
    select(theme, vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
    kable(caption = sprintf(
      "One-vs-rest favorability score test by theme (%s, %s axis)",
      rendering, axis_name
    )) %>%
    kable_styling(full_width = FALSE,
                  bootstrap_options = c("striped", "hover"))
}
```

## Clinical utility by theme

```{r favorability_clinical_by_theme, results='asis'}
themes <- sort(unique(qa_ratings$primary_question_type))

for (theme in themes) {
  cat(sprintf("\n\n### %s\n\n", theme))
  results <- run_favorability_analysis_by_theme(qa_ratings, "clinical_utility", theme)
  if (!is.null(results) && nrow(results) > 0) {
    print(
      results %>%
        mutate(
          favorability = sprintf("%.1f%%", 100 * favorability),
          ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
          p_value = sprintf("%.4f", p_value)
        ) %>%
        select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
        kable(caption = sprintf("Clinical utility: %s", theme)) %>%
        kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
    )
  } else {
    cat("No data available for this theme.\n")
  }
}
```

## Accuracy by theme

```{r favorability_accuracy_by_theme, results='asis'}
for (theme in themes) {
  cat(sprintf("\n\n### %s\n\n", theme))
  results <- run_favorability_analysis_by_theme(qa_ratings, "accuracy", theme)
  if (!is.null(results) && nrow(results) > 0) {
    print(
      results %>%
        mutate(
          favorability = sprintf("%.1f%%", 100 * favorability),
          ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
          p_value = sprintf("%.4f", p_value)
        ) %>%
        select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
        kable(caption = sprintf("Accuracy: %s", theme)) %>%
        kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
    )
  } else {
    cat("No data available for this theme.\n")
  }
}
```

## Authoritativeness by theme

```{r favorability_authoritativeness_by_theme, results='asis'}
for (theme in themes) {
  cat(sprintf("\n\n### %s\n\n", theme))
  results <- run_favorability_analysis_by_theme(qa_ratings, "authoritativeness", theme)
  if (!is.null(results) && nrow(results) > 0) {
    print(
      results %>%
        mutate(
          favorability = sprintf("%.1f%%", 100 * favorability),
          ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
          p_value = sprintf("%.4f", p_value)
        ) %>%
        select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
        kable(caption = sprintf("Authoritativeness: %s", theme)) %>%
        kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
    )
  } else {
    cat("No data available for this theme.\n")
  }
}
```

## Completeness by theme

```{r favorability_completeness_by_theme, results='asis'}
for (theme in themes) {
  cat(sprintf("\n\n### %s\n\n", theme))
  results <- run_favorability_analysis_by_theme(qa_ratings, "completeness", theme)
  if (!is.null(results) && nrow(results) > 0) {
    print(
      results %>%
        mutate(
          favorability = sprintf("%.1f%%", 100 * favorability),
          ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
          p_value = sprintf("%.4f", p_value)
        ) %>%
        select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
        kable(caption = sprintf("Completeness: %s", theme)) %>%
        kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
    )
  } else {
    cat("No data available for this theme.\n")
  }
}
```

## Verifiability by theme

```{r favorability_verifiability_by_theme, results='asis'}
for (theme in themes) {
  cat(sprintf("\n\n### %s\n\n", theme))
  results <- run_favorability_analysis_by_theme(qa_ratings, "verifiability", theme)
  if (!is.null(results) && nrow(results) > 0) {
    print(
      results %>%
        mutate(
          favorability = sprintf("%.1f%%", 100 * favorability),
          ci_95 = sprintf("[%.1f%%, %.1f%%]", 100 * ci_lower, 100 * ci_upper),
          p_value = sprintf("%.4f", p_value)
        ) %>%
        select(vendor, wins, ties, losses, n, favorability, ci_95, p_value) %>%
        kable(caption = sprintf("Verifiability: %s", theme)) %>%
        kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
    )
  } else {
    cat("No data available for this theme.\n")
  }
}
```

# Save outputs

```{r save_outputs, eval=FALSE}
# output_dir <- ifelse(is.null(params$output_dir), params$data_dir, params$output_dir)
# rdata_path <- file.path(output_dir, paste0(params$output_name, ".RData"))
# save.image(file = rdata_path)
# message("Saved RData to: ", rdata_path)

# results_list <- list(
#   clinical_by_theme = all_clinical_results,
#   accuracy_by_theme = all_accuracy_results,
#   theme_distribution = theme_counts
# )
# rds_path <- file.path(output_dir, paste0(params$output_name, "_results.rds"))
# saveRDS(results_list, file = rds_path)
# message("Saved results to: ", rds_path)
```
