---
title: "LLM-as-a-Judge Analysis"
author: "Jean Feng"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  data_dir: null
  output_dir: null
  output_name: "judge_analysis"
  bootstrap_by: "question"
  n_boot: 0
  n_boot_corr: 1000
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(jsonlite)
library(knitr)
library(kableExtra)

source("favorability_test.R")
source("data_helpers.R")
```

# Analysis parameters

```{r params, echo=FALSE, results='asis'}
cat(sprintf("- **Data directory:** `%s`\n", params$data_dir))
cat(sprintf("- **Output directory:** `%s`\n", ifelse(is.null(params$output_dir), params$data_dir, params$output_dir)))
cat(sprintf("- **Bootstrap by:** %s\n", params$bootstrap_by))
cat(sprintf("- **N bootstrap (favorability):** %d\n", params$n_boot))
cat(sprintf("- **N bootstrap (correlation CI):** %d\n", params$n_boot_corr))
```

# Load human ratings data

Replicate the data loading from analysis.Rmd.

```{r load_human}
if (is.null(params$data_dir)) stop("`data_dir` parameter is required")
data_dir <- params$data_dir

assignments <- read_csv(file.path(data_dir, "blinded_qa_assignments.csv"))
questions   <- read_csv(file.path(data_dir, "blinded_qa_questions.csv"))
ratings     <- read_csv(file.path(data_dir, "blinded_ratings.csv"))
user_registrations  <- read_csv(file.path(data_dir, "recruitment_combined.csv"))
users       <- read_csv(file.path(data_dir, "blinded_users.csv"))
users <- get_eligible_users(users, user_registrations, user_stratification = NULL)
prior_users_path <- file.path(data_dir, "blinded_prior_users.csv")
prior_users <- if (file.exists(prior_users_path)) read_csv(prior_users_path) else NULL

qa_ratings_human <- build_qa_ratings(ratings, assignments, questions, users, prior_users = prior_users)
qa_ratings_human <- apply_inclusion_criteria(qa_ratings_human)
```

# Load vendor blinding mapping

```{r load_mapping}
secret_mapping <- read_csv(file.path(data_dir, "secret_mapping.csv"))
provider_to_vendor <- setNames(secret_mapping$vendor, secret_mapping$provider)
```

# Load and process machine ratings

Machine ratings are joined to assignments via `assignmentId` to get the
question_id and blinded slot vendors.

```{r load_machine}
machine_raw <- stream_in(file(file.path(data_dir, "machineRatings.jsonl")), verbose = FALSE)

choices_df <- machine_raw$choices
choices_df$row_id <- seq_len(nrow(choices_df))

machine_long <- machine_raw %>%
  mutate(row_id = row_number()) %>%
  select(row_id, assignment_id = assignmentId, grader, variant, subversion) %>%
  left_join(
    choices_df %>%
      pivot_longer(cols = -row_id, names_to = "axis", values_to = "choice"),
    by = "row_id"
  ) %>%
  select(-row_id)

machine_long <- machine_long %>%
  inner_join(
    assignments %>% select(assignment_id = id, question_id, slot_a_vendor, slot_b_vendor),
    by = "assignment_id"
  ) %>%
  mutate(grader_vendor = provider_to_vendor[grader])

machine_long <- machine_long %>%
  mutate(
    winner = case_when(
      choice %in% c("strongly_a", "slightly_a") ~ slot_a_vendor,
      choice %in% c("strongly_b", "slightly_b") ~ slot_b_vendor,
      choice == "tie" ~ NA_character_
    ),
    loser = case_when(
      choice %in% c("strongly_a", "slightly_a") ~ slot_b_vendor,
      choice %in% c("strongly_b", "slightly_b") ~ slot_a_vendor,
      choice == "tie" ~ NA_character_
    ),
    is_tie = choice == "tie"
  )

cat(sprintf("Total machine ratings: %d\n", nrow(machine_long)))
cat(sprintf("Unique graders: %s\n", paste(unique(machine_long$grader), collapse = ", ")))
```

# Filter to text-only rendering (matching primary analysis)

The primary analysis uses `qa_text_only` subversion. For machine judges, we use
`variant == "text_only"` to match. We also restrict to the same subversion.

```{r filter_text_only}
n_na_choice <- sum(is.na(machine_long$choice))
cat(sprintf("Dropping %d ratings with NA choice\n", n_na_choice))

machine_text_only <- machine_long %>%
  filter(variant == "text_only", subversion == "qa_text_only", !is.na(choice))

cat(sprintf("Machine ratings after filtering to text_only/qa_text_only: %d\n", nrow(machine_text_only)))
```

# Compute human favorability scores (reference)

```{r human_favorability_helpers}
prepare_one_vs_rest_df <- function(df, vendor) {
  df %>%
    filter(slot_a_vendor == vendor | slot_b_vendor == vendor) %>%
    mutate(
      outcome = case_when(
        winner == vendor ~ 1L,
        loser == vendor ~ -1L,
        is_tie ~ 0L
      )
    ) %>%
    select(question_id, outcome)
}

run_favorability_analysis_simple <- function(df, axis_name, n_boot = 0,
                                             seed = 42, bootstrap_by = "question") {
  set.seed(seed)
  axis_df <- df %>% filter(axis == axis_name)
  if (nrow(axis_df) == 0) return(tibble())

  vendors <- sort(unique(c(axis_df$slot_a_vendor, axis_df$slot_b_vendor)))

  map_dfr(vendors, function(v) {
    test_df <- prepare_one_vs_rest_df(axis_df, v)
    if (nrow(test_df) == 0) return(NULL)

    fav_score <- get_favorability_score(test_df$outcome)

    if (n_boot > 0) {
      test_df$user_id <- seq_len(nrow(test_df))
      test_result <- run_favorability_test(test_df, n_boot = n_boot, ci_level = NULL,
                                           bootstrap_by = bootstrap_by)
      ci_lower <- test_result$ci_lower
      ci_upper <- test_result$ci_upper
    } else {
      ci_lower <- NA_real_
      ci_upper <- NA_real_
    }

    tibble(
      vendor = v,
      wins = sum(test_df$outcome == 1, na.rm = TRUE),
      ties = sum(test_df$outcome == 0, na.rm = TRUE),
      losses = sum(test_df$outcome == -1, na.rm = TRUE),
      n = nrow(test_df),
      favorability = fav_score,
      ci_lower = ci_lower,
      ci_upper = ci_upper
    )
  })
}
```

```{r human_favorability}
human_text_only <- qa_ratings_human %>% filter(subversion == "qa_text_only")
axes <- c("accuracy", "clinical_utility", "authoritativeness", "completeness", "verifiability")

human_favorability <- map_dfr(axes, function(ax) {
  run_favorability_analysis_simple(human_text_only, ax, n_boot = params$n_boot,
                                   bootstrap_by = params$bootstrap_by) %>%
    mutate(axis = ax, judge = "human")
})

human_favorability %>%
  select(axis, vendor, favorability, ci_lower, ci_upper) %>%
  mutate(
    favorability_fmt = sprintf("%.1f%% [%.1f%%, %.1f%%]",
                               100 * favorability, 100 * ci_lower, 100 * ci_upper)
  ) %>%
  select(axis, vendor, favorability_fmt) %>%
  pivot_wider(names_from = axis, values_from = favorability_fmt) %>%
  kable(caption = "Human expert favorability scores (reference)") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Compute judge favorability scores

For each LLM judge, compute one-vs-rest favorability scores.

```{r judge_favorability}
graders <- unique(machine_text_only$grader)

judge_favorability <- map_dfr(graders, function(g) {
  judge_df <- machine_text_only %>% filter(grader == g)

  map_dfr(axes, function(ax) {
    run_favorability_analysis_simple(judge_df, ax, n_boot = params$n_boot,
                                     bootstrap_by = params$bootstrap_by) %>%
      mutate(axis = ax, judge = g)
  })
})
```

# Compute jury favorability scores

The jury aggregates ratings by averaging Likert scores across judges, then
determining win/tie/lose based on the average.

```{r jury_favorability}
choice_to_score <- c(strongly_a = -2, slightly_a = -1, tie = 0, slightly_b = 1, strongly_b = 2)

jury_ratings <- machine_text_only %>%
  mutate(score = choice_to_score[choice]) %>%
  group_by(question_id, slot_a_vendor, slot_b_vendor, axis) %>%
  summarise(
    avg_score = mean(score, na.rm = TRUE),
    n_judges = n(),
    .groups = "drop"
  ) %>%
  mutate(
    winner = case_when(
      avg_score < -0.5 ~ slot_a_vendor,
      avg_score > 0.5 ~ slot_b_vendor,
      TRUE ~ NA_character_
    ),
    loser = case_when(
      avg_score < -0.5 ~ slot_b_vendor,
      avg_score > 0.5 ~ slot_a_vendor,
      TRUE ~ NA_character_
    ),
    is_tie = abs(avg_score) <= 0.5
  )

jury_favorability <- map_dfr(axes, function(ax) {
  run_favorability_analysis_simple(jury_ratings, ax, n_boot = params$n_boot,
                                   bootstrap_by = params$bootstrap_by) %>%
    mutate(axis = ax, judge = "jury")
})
```

# Combine all favorability results

```{r combine_favorability}
all_favorability <- bind_rows(
  human_favorability,
  judge_favorability,
  jury_favorability
)

all_favorability %>%
  select(judge, axis, vendor, favorability) %>%
  mutate(favorability = round(100 * favorability, 1)) %>%
  pivot_wider(names_from = vendor, values_from = favorability) %>%
  arrange(axis, judge) %>%
  kable(caption = "One-vs-rest favorability (%) by judge and axis") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Compute Kendall's tau correlation: judge vs human

For each judge (and jury), compute Kendall's tau correlation between the
judge's favorability scores and the human expert's scores. We use Kendall's tau
since we only have 4 vendors to rank. We compute per-axis correlations and then
average across axes, bootstrapping to get meaningful CIs.

```{r kendall_correlation}
compute_kendall_vs_human <- function(fav_df) {
  human_df <- fav_df %>% filter(judge == "human") %>% select(axis, vendor, human_fav = favorability)

  fav_df %>%
    filter(judge != "human") %>%
    left_join(human_df, by = c("axis", "vendor")) %>%
    group_by(judge, axis) %>%
    summarise(
      kendall = cor(favorability, human_fav, method = "kendall"),
      n_vendors = n(),
      .groups = "drop"
    )
}

compute_avg_kendall_vs_human <- function(fav_df) {
  per_axis <- compute_kendall_vs_human(fav_df)
  per_axis %>%
    group_by(judge) %>%
    summarise(
      avg_kendall = mean(kendall, na.rm = TRUE),
      n_axes = sum(!is.na(kendall)),
      .groups = "drop"
    )
}

kendall_per_axis <- compute_kendall_vs_human(all_favorability)
kendall_avg <- compute_avg_kendall_vs_human(all_favorability)

kendall_per_axis %>%
  pivot_wider(names_from = axis, values_from = kendall, names_prefix = "tau_") %>%
  kable(caption = "Per-axis Kendall's tau of judge favorability vs human expert (n=4 vendors each)",
        digits = 3) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Bootstrap confidence intervals for average Kendall's tau

Bootstrap by question to compute 95% CI for the average Kendall's tau
across all axes.

```{r bootstrap_kendall}
bootstrap_avg_kendall <- function(human_df, machine_df, axes,
                                   n_boot = 1000, seed = 42,
                                   bootstrap_by = "question") {
  set.seed(seed)

  graders <- unique(machine_df$grader)
  question_ids <- unique(c(human_df$question_id, machine_df$question_id))

  boot_results <- vector("list", n_boot)

  for (b in seq_len(n_boot)) {
    boot_questions <- sample(question_ids, length(question_ids), replace = TRUE)

    # Create bootstrap instances with unique IDs to prevent downstream collapse
    boot_instances <- tibble(question_id = boot_questions) %>%
      group_by(question_id) %>%
      mutate(boot_instance = row_number()) %>%
      ungroup()

    # Join and assign unique question_id per bootstrap instance
    human_boot <- human_df %>%
      inner_join(boot_instances, by = "question_id", relationship = "many-to-many") %>%
      mutate(question_id = paste0(question_id, "_", boot_instance)) %>%
      select(-boot_instance)
    machine_boot <- machine_df %>%
      inner_join(boot_instances, by = "question_id", relationship = "many-to-many") %>%
      mutate(question_id = paste0(question_id, "_", boot_instance)) %>%
      select(-boot_instance)

    jury_boot <- machine_boot %>%
      mutate(score = choice_to_score[choice]) %>%
      group_by(question_id, slot_a_vendor, slot_b_vendor, axis) %>%
      summarise(avg_score = mean(score, na.rm = TRUE), .groups = "drop") %>%
      mutate(
        winner = case_when(avg_score < -0.5 ~ slot_a_vendor, avg_score > 0.5 ~ slot_b_vendor, TRUE ~ NA_character_),
        loser = case_when(avg_score < -0.5 ~ slot_b_vendor, avg_score > 0.5 ~ slot_a_vendor, TRUE ~ NA_character_),
        is_tie = abs(avg_score) <= 0.5
      )

    boot_fav <- map_dfr(axes, function(ax) {
      human_fav <- run_favorability_analysis_simple(human_boot, ax, n_boot = 0, bootstrap_by = bootstrap_by)
      if (is.null(human_fav) || nrow(human_fav) == 0) return(NULL)
      human_fav <- human_fav %>% mutate(axis = ax, judge = "human")

      judge_fav <- map_dfr(graders, function(g) {
        judge_df <- machine_boot %>% filter(grader == g)
        res <- run_favorability_analysis_simple(judge_df, ax, n_boot = 0, bootstrap_by = bootstrap_by)
        if (is.null(res) || nrow(res) == 0) return(NULL)
        res %>% mutate(axis = ax, judge = g)
      })

      jury_fav <- run_favorability_analysis_simple(jury_boot, ax, n_boot = 0, bootstrap_by = bootstrap_by)
      if (!is.null(jury_fav) && nrow(jury_fav) > 0) {
        jury_fav <- jury_fav %>% mutate(axis = ax, judge = "jury")
      } else {
        jury_fav <- NULL
      }

      bind_rows(human_fav, judge_fav, jury_fav)
    })

    if (nrow(boot_fav) == 0) next

    boot_avg_kendall <- compute_avg_kendall_vs_human(boot_fav)
    boot_avg_kendall$boot_iter <- b
    boot_results[[b]] <- boot_avg_kendall
  }

  bind_rows(boot_results)
}

boot_kendall_df <- bootstrap_avg_kendall(
  human_text_only, machine_text_only, axes,
  n_boot = params$n_boot_corr, seed = 42, bootstrap_by = params$bootstrap_by
)
```

```{r kendall_ci}
# Pivotal bootstrap CI: [2*theta_hat - q_upper, 2*theta_hat - q_lower]
kendall_ci <- boot_kendall_df %>%
  group_by(judge) %>%
  summarise(
    q_lower = quantile(avg_kendall, 0.025, na.rm = TRUE),
    q_upper = quantile(avg_kendall, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

kendall_final <- kendall_avg %>%
  left_join(kendall_ci, by = "judge") %>%
  mutate(
    ci_lower = 2 * avg_kendall - q_upper,
    ci_upper = 2 * avg_kendall - q_lower,
    kendall_fmt = sprintf("%.3f [%.3f, %.3f]", avg_kendall, ci_lower, ci_upper)
  ) %>%
  arrange(desc(avg_kendall))

kendall_final %>%
  select(judge, kendall_fmt) %>%
  kable(caption = "Average Kendall's tau vs human expert (across 5 axes) with 95% pivotal bootstrap CI",
        col.names = c("Judge", "Average Kendall's tau [95% CI]")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

# Bootstrap correlation histograms

Visualize the distribution of bootstrap correlations for each judge to diagnose
the bootstrap confidence intervals.

```{r bootstrap_histograms, fig.width=10, fig.height=8}
boot_kendall_df %>%
  left_join(kendall_final %>% select(judge, point_est = avg_kendall, ci_lower, ci_upper), by = "judge") %>%
  ggplot(aes(x = avg_kendall)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white", alpha = 0.7) +
  geom_vline(aes(xintercept = point_est), color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(aes(xintercept = ci_lower), color = "darkgreen", linetype = "dotted", linewidth = 0.8) +
  geom_vline(aes(xintercept = ci_upper), color = "darkgreen", linetype = "dotted", linewidth = 0.8) +
  facet_wrap(~ judge, scales = "free_y", ncol = 2) +
  labs(
    title = "Bootstrap distribution of average Kendall's tau (judge vs human)",
    subtitle = "Red dashed = point estimate; Green dotted = 95% pivotal CI bounds",
    x = "Average Kendall's tau",
    y = "Count"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold", size = 11))
```


# Save outputs

```{r save_outputs}
output_dir <- ifelse(is.null(params$output_dir), params$data_dir, params$output_dir)

rdata_path <- file.path(output_dir, paste0(params$output_name, ".RData"))
save.image(file = rdata_path)
message("Saved RData to: ", rdata_path)

write_csv(all_favorability, file.path(output_dir, "judge_favorability.csv"))
write_csv(kendall_per_axis, file.path(output_dir, "judge_kendall_per_axis.csv"))
write_csv(kendall_final, file.path(output_dir, "judge_kendall_avg.csv"))
message("Saved CSV outputs to: ", output_dir)
```
